---
title: "Preliminary Mediator Associations Report"
subtitle: "Detailed Methods and Findings from Dataset Preparation"
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false
read_obj <- function(filename) {
  candidates <- c(
    file.path("..", "_tools", "_objects", filename),
    file.path("_tools", "_objects", filename)
  )
  idx <- which(file.exists(candidates))[1]
  if (is.na(idx)) {
    stop("Missing object file: ", filename)
  }
  readRDS(candidates[idx])
}

summary_df <- read_obj("mediation_prelim_summary.rds")
pair_assoc_df <- read_obj("mediation_prelim_pair_assoc.rds")
interpretation <- tryCatch(
  read_obj("mediation_prelim_interpretation.rds"),
  error = function(e) NULL
)
build_info <- read_obj("mediation_prelim_build_info.rds")
combined_plot_path <- file.path("..", "_tools", "_objects", "mediation_prelim_combined_plot.png")

fmt_p <- function(x) {
  ifelse(is.na(x), NA_character_, format.pval(x, digits = 3, eps = 0.001))
}
```

## Scope

This report reflects the detailed analytical intent and implementation in the dataset preparation source index, specifically:

1.  `Mediators vs Outcome`.
2.  `Potential Mediators: Inter-association and Correlation`.

This web report presents outcome-association and pairwise inter-association outputs using aggregated objects only (no row-level data stored in this project).

## Source Alignment (Dataset Preparation)

The report mirrors the dataset preparation source workflow:

1.  Recode and harmonize analysis variables from `dt_multistudies`.
2.  Fit three logistic models for mediator-to-outcome association.
3.  Quantify pairwise mediator dependence with OR and Cramer's V.
4.  Provide interpretation anchored to fully adjusted and pairwise strength patterns.

## Analytical Inputs

```{r}
#| echo: false
input_tbl <- data.frame(
  Component = c(
    "Outcome",
    "Candidate mediators",
    "Covariates in base/fully-adjusted models",
    "Key preprocessing from source workflow",
    "Data source at build time"
  ),
  Specification = c(
    "DPBN (binary depression indicator)",
    "DNKA, ESXC, FDSC, GOVG, VLNC, SCHL, SCSP",
    "SEX, AGE, HIVS, VSTY, EXTMG, RURAL, ORPH",
    "GOVG recoded to 'No Government Grant'; VSTY derived from VISITDT year and centered.",
    build_info$source
  ),
  check.names = FALSE
)

knitr::kable(input_tbl, align = c("l", "l"), row.names = FALSE)
```

## Model Specifications (from source workflow)

```{r}
#| echo: false
model_tbl <- data.frame(
  Model = c(
    "Unadjusted",
    "Base model + single mediator",
    "Fully adjusted model"
  ),
  Formula = c(
    "DPBN ~ mediator",
    "DPBN ~ mediator + SEX + AGE + HIVS + VSTY + EXTMG + RURAL + ORPH",
    "DPBN ~ DNKA + ESXC + FDSC + GOVG + VLNC + SCHL + SCSP + SEX + AGE + HIVS + VSTY + EXTMG + RURAL (+ ORPH as specified in source build)"
  ),
  Interpretation = c(
    "Crude association per mediator.",
    "Mediator association adjusted for prespecified confounding set.",
    "Independent mediator signal with all candidate mediators entered simultaneously."
  ),
  check.names = FALSE
)

knitr::kable(model_tbl, align = c("l", "l", "l"), row.names = FALSE)
```

## Build Metadata

```{r}
#| echo: false
meta_tbl <- data.frame(
  Field = c("Rendered at", "Analysis rows", "Mediator pairs"),
  Value = c(
    build_info$rendered_at,
    build_info$n_analysis,
    build_info$n_pairs
  ),
  check.names = FALSE
)

knitr::kable(meta_tbl, align = c("l", "l"), row.names = FALSE)
```

## Outcome Associations (Unadjusted, Base, Fully Adjusted)

```{r}
#| echo: false
outcome_tbl <- data.frame(
  `Potential Mediator` = summary_df$label,
  `Unadjusted OR (95% CI)` = summary_df$unadj_or_ci,
  `Unadjusted p-value` = fmt_p(summary_df$unadj_p.value),
  `Base OR (95% CI)` = summary_df$base_or_ci,
  `Base p-value` = fmt_p(summary_df$base_p.value),
  `Fully Adjusted OR (95% CI)` = summary_df$adj_or_ci,
  `Fully Adjusted p-value` = fmt_p(summary_df$adj_p.value),
  check.names = FALSE
)

knitr::kable(outcome_tbl, row.names = FALSE)
```

```{r}
#| echo: false
knitr::include_graphics(combined_plot_path)
```

## Pairwise Inter-association (OR and Cramer's V)

```{r}
#| echo: false
pair_tbl <- data.frame(
  `Mediator A` = pair_assoc_df$label_a,
  `Mediator B` = pair_assoc_df$label_b,
  `Complete N` = pair_assoc_df$n_complete,
  `OR (95% CI)` = ifelse(
    is.na(pair_assoc_df$or),
    "NA",
    sprintf("%.2f (%.2f, %.2f)", pair_assoc_df$or, pair_assoc_df$conf.low, pair_assoc_df$conf.high)
  ),
  `OR p-value` = fmt_p(pair_assoc_df$p_value),
  `Cramer's V` = round(pair_assoc_df$cramer_v, 3),
  check.names = FALSE
)
ord <- order(-pair_tbl$`Cramer's V`, pair_tbl$`OR p-value`)
pair_tbl <- pair_tbl[ord, , drop = FALSE]

knitr::kable(pair_tbl, row.names = FALSE)
```

## Interpretation

### Outcome Association Interpretation

```{r}
#| echo: false
if (!is.null(interpretation) && length(interpretation) > 0) {
  cat(paste0("- ", interpretation), sep = "\n")
} else {
  sig_adj <- summary_df[!is.na(summary_df$adj_p.value) & summary_df$adj_p.value < 0.05, , drop = FALSE]
  if (nrow(sig_adj) > 0) {
    sig_adj <- sig_adj[order(-sig_adj$adj_estimate), , drop = FALSE]
    sig_adj <- head(sig_adj, 3)
  }

  if (nrow(sig_adj) == 0) {
    cat("- No mediator reached statistical significance in the fully adjusted model.\n")
  } else {
    txt <- paste(
      sig_adj$label,
      sprintf("(aOR %s)", sig_adj$adj_or_ci),
      collapse = "; "
    )
    cat("- Strongest fully adjusted signals are: \n", txt, ".\n", sep = "")
  }
}
cat("- Unadjusted estimates are descriptive; fully adjusted estimates should guide prioritization.\n")
```

### Pairwise Interpretation

```{r}
#| echo: false
pair_interpret <- pair_assoc_df[!is.na(pair_assoc_df$cramer_v), , drop = FALSE]
if (nrow(pair_interpret) > 0) {
  pair_interpret <- pair_interpret[
    order(-pair_interpret$cramer_v, pair_interpret$p_value),
    ,
    drop = FALSE
  ]
  pair_interpret <- head(pair_interpret, 3)
}

if (nrow(pair_interpret) == 0) {
  cat("- Pairwise inter-association summary was not available for interpretation.\n")
} else {
  for (i in seq_len(nrow(pair_interpret))) {
    r <- pair_interpret[i, ]
    cat(
      "- ",
      r$label_a,
      " vs ",
      r$label_b,
      ": OR ",
      ifelse(is.na(r$or), "NA", sprintf("%.2f", r$or)),
      ", Cramer's V ",
      ifelse(is.na(r$cramer_v), "NA", sprintf("%.2f", r$cramer_v)),
      ".\n",
      sep = ""
    )
  }
}
cat("- Inter-association identifies co-patterning among mediators and does not imply causal direction.\n")
```

## Correlation Reporting Decision

Correlation heatmap and rho columns are intentionally omitted in this report version.\
For binary 2x2 mediator pairs, the correlation measure (phi/rho in this coding) is numerically equivalent to Cramer's V; therefore, Cramer's V is retained as the single strength metric.
