---
title: "Data Preparation and Descriptive Analysis"
subtitle: "**Investigating Causal Pathways from Early-Life Exposure Trajectories to Depression in Adolescence**"
date: today
format:
  html: 
    grid:
      sidebar-width: 300px
      body-width: 1200px
      margin-width: 300px
      gutter-width: 1.5rem
    theme: sandstone
    embed-resources: true
    self-contained: true
    page-layout: full
    fig-dpi: 300
    toc: true
    number-sections: true
    code-fold: true
    
  docx: 
    toc: true
    
  pdf: 
    toc: true
editor: visual
---

## Setup

This section loads the necessary R packages for the analysis. The `tidyverse` package will be used for general data manipulation, `here` for robust file path management, `gtsummary` for creating descriptive tables, `naniar` for visualizing missing data, and `ggsankey` for the Sankey plot.

```{r setup, message=FALSE, warning=FALSE}
# Load required libraries
library(tidyverse)
library(here)
library(gtsummary)
library(naniar)
library(ggsankey)
library(ComplexUpset)
library(patchwork)
library(gt)
library(gtreg)
library(gtsummary)

source(here("../../pipelines_and_functions/data_processing_and_description_funcs.R"))

# Global figure size defaults for all plots
knitr::opts_chunk$set(fig.width = 12, fig.height = 5)
```

## Load Datasets

The analysis utilizes two primary datasets: `dt_childhood_exposure` and `dt_multistudies`. These files are stored in `.RData` format. We will use the `load()` function, which loads the R objects directly into the global environment. The `here()` function is used to construct a reliable file path from the project's root directory.

```{r load-data}
# Define the path to the data directory
data_path <- here("statistical_analysis", "data", "adam")

# Load the childhood exposure dataset
load(file.path(data_path, "dt_childhood_exposure.RData"))

# Load the adolescent multi-study dataset
load(file.path(data_path, "dt_multistudies.RData"))

# Define the output path
output_path <- here("statistical_analysis", "data", "adam")
```

### Initial Data Inspection

After loading, it is good practice to inspect the structure of the datasets using `glimpse()`. This allows us to ensure they have been loaded correctly, review the variable names and types, and see the number of observations.

```{r inspect-data}
# Glimpse the childhood exposure data
glimpse(dt_childhood_exposure)

# Glimpse the adolescent data
glimpse(dt_multistudies)
```

## Prepare Childhood Exposure Data

This section prepares the `dt_childhood_exposure` dataset for analysis. The primary goal is to ensure the dataset aligns with the study population defined in the SAP, which is anchored by participants present in the adolescent `dt_multistudies` data.

### 1. Identify the Analysis Cohort

As specified in the SAP, the final study population is defined by participants who have data in the adolescent wave (`dt_multistudies`). The following code chunk extracts a unique list of participant identifiers (`USUBJID`) from the adolescent (Age between 13 and 24) dataset to serve as our definitive analysis cohort.

```{r identify-cohort}
# Get unique participant IDs from the adolescent data
analysis_cohort_ids <- dt_multistudies %>% 
  filter(AGE>=13 & AGE<=24) |> 
  distinct(USUBJID)
```

The final analysis cohort consists of `r nrow(analysis_cohort_ids)` unique participants, identified from the adolescent dataset.

### 2. Filter Childhood Data to Match Cohort

To ensure consistency, we must subset the longitudinal childhood data to include only the participants who are part of our final study cohort. This code filters the `dt_childhood_exposure` dataframe, retaining rows where the `USUBJID` is present in the `analysis_cohort_ids` list created above.

```{r filter-childhood-data}
# Filter the childhood data
childhood_filtered_cohort <- dt_childhood_exposure %>%
  filter(USUBJID %in% analysis_cohort_ids$USUBJID) 
```

After filtering the childhood data to match this cohort, `r n_distinct(childhood_filtered_cohort$USUBJID)` participants and `r nrow(childhood_filtered_cohort)` total observations remain.

### 3. Check Data Integrity and Apply Age Filter

A critical step is to check for data quality issues and enforce the age range for the exposure period. First, a data integrity check is performed to identify any instances where a participant has multiple records for the same exposure year (`EXPAGE`). Second, the data is filtered to retain only records where `EXPAGE` is between 0 and 10, as specified in the SAP.

```{r check-integrity}
# Check for duplicates: count rows per participant per exposure year
duplicate_check <- childhood_filtered_cohort %>%
  group_by(USUBJID, EXPAGE) %>%
  summarise(n = n(), .groups = 'drop') %>%
  filter(n > 1)

# Report if any duplicates were found
if (nrow(duplicate_check) > 0) {
  print(duplicate_check)
} else {
  print("Data Integrity Check: No duplicate participant-year entries found.")
}

# Filter for the relevant exposure period (ages 0-10)
childhood_filtered <- childhood_filtered_cohort %>%
  filter(EXPAGE >= 0 & EXPAGE <= 10)

# Check the range of the exposure year variable after filtering
summary(childhood_filtered$EXPAGE)

## Variable Selected Variables
childhood_data <- dt_childhood_exposure %>%
  dplyr::select(USUBJID, EXPAGE, LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4)
save(childhood_data, file = file.path(output_path, "childhood_data.RData"))


```

The summary above shows the range of `EXPAGE` after filtering to the required 0-10 year period. The final childhood dataset ready for transformation contains `r n_distinct(childhood_filtered$USUBJID)` unique participants and `r nrow(childhood_filtered)` total observations.

## Datasets Metadata

```{r metadata}
func_dictionary(childhood_data) |> 
  gt()


func_dictionary(dt_multistudies) |> 
  gt()
```

## Participant Data Availability Flow

To understand the overlap in data availability between the two key developmental periods, we can visualize the flow of participants using a Sankey diagram. This plot shows how many participants have data in the early-life period (0-5 years), the middle-childhood period (6-10 years), or both.

```{r sankey-plot}
# Participant Data Availability Flow between Cohort and Availability (reverted + right-side labels with % of total cohort)

# Determine data availability for each participant
data_availability <- childhood_filtered %>%
  group_by(USUBJID) %>%
  summarise(
    has_0_5 = any(EXPAGE >= 0 & EXPAGE <= 5),
    has_6_10 = any(EXPAGE >= 6 & EXPAGE <= 10)
  ) %>%
  mutate(
    Availability = case_when(
      has_0_5 & has_6_10  ~ "Data for Both Periods",
      has_0_5 & !has_6_10 ~ "Data for 0-5 Years Only",
      !has_0_5 & has_6_10 ~ "Data for 6-10 Years Only",
      TRUE ~ "No Data in Range"
    )
  ) %>%
  filter(Availability != "No Data in Range") %>%
  mutate(Cohort = "Total Childhood Cohort")

# Convert to long format for ggsankey (Cohort -> Availability)
sankey_long <- ggsankey::make_long(data_availability, Cohort, Availability)

# Precompute node counts per stage for labels (avoid after_stat()/n())
sankey_counts <- sankey_long %>%
  dplyr::count(x, node, name = "n_lab")

# Total cohort size for percentage denominator
total_cohort <- nrow(analysis_cohort_ids)

# Join counts back and compute percent of total cohort
sankey_long <- sankey_long %>%
  dplyr::left_join(sankey_counts, by = c("x", "node")) %>%
  dplyr::mutate(p_lab = 100 * n_lab / total_cohort)

# Generate the Sankey plot with labels to the right of nodes, including n and % of total
p1<-ggplot(sankey_long, aes(x = x,
                       next_x = next_x,
                       node = node,
                       next_node = next_node,
                       fill = factor(node))) +
  geom_sankey(flow.alpha = 0.6) +
  geom_sankey_label(
    aes(label = paste0(node, "
", n_lab, " (", sprintf("%.1f%%", p_lab), ")")),
    size = 3.5, color = "white", fill = "gray40", hjust = 0,
    position = position_nudge(x = 0.05)
  ) +
  scale_fill_viridis_d(guide = "none") +
  theme_minimal(base_size = 16) +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none") +
  labs(x = NULL,
       title = "Participant Data Availability Flow",
       subtitle = "Labels show count and % of the total cohort; labels nudged to the right of each node")
```

```{r}
DTUPSET<-dt_multistudies |> 
  dplyr::select(USUBJID,STUDY) |> 
  group_by(USUBJID) |> 
  mutate(FLAG=TRUE) |> 
  distinct() |> 
  spread(STUDY,FLAG) |> 
  mutate(across(where(is.logical),~ifelse(is.na(.x),FALSE,.x)))

studies<-names(DTUPSET)[-1]

set_size = function(w, h, factor=1.5) {
    s = 1 * factor
    options(
        repr.plot.width=w * s,
        repr.plot.height=h * s,
        repr.plot.res=100 / factor,
        jupyter.plot_mimetypes='image/png',
        jupyter.plot_scale=1
    )
}
set_size(8, 3)
p2<-ComplexUpset::upset(DTUPSET,
      studies)
```


```{r, fig.height=10,fig.width=12}
p1/p2
```

## Visualize Missing Data Patterns

Visualizing the pattern of missing data is crucial for understanding potential biases and informing the multiple imputation strategy outlined in the SAP. The following plot shows the percentage of complete (non-missing) observations for each RMLCA indicator variable across the full childhood period (ages 0-10).

```{r visualize-missingness}
# Calculate the percentage of complete data per year for each indicator
missingness_summary <- childhood_filtered %>%
  group_by(EXPAGE) %>%
  summarise(across(
    .cols = c(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4),
    .fns = ~ (1 - sum(is.na(.)) / n()) * 100, # Calculate percentage complete
    .names = "{.col}_complete"
  )) %>%
  pivot_longer(
    cols = -EXPAGE,
    names_to = "Indicator",
    values_to = "PercentComplete",
    names_pattern = "(.*)_complete"
  ) |> 
  mutate(Data="Original")

# Calculate the percentage of complete data per year for each indicator
missingness_summary_filled <- childhood_filtered %>%
  group_by(USUBJID) |> 
  arrange(USUBJID,EXPAGE) |> 
  fill(EDUCMOTH, EDUCFATH, HHSES, .direction = "down") |> 
  fill(HHCHLD4, .direction = "up") |> 
  group_by(EXPAGE) %>%
  summarise(across(
    .cols = c(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4),
    .fns = ~ (1 - sum(is.na(.)) / n()) * 100, # Calculate percentage complete
    .names = "{.col}_complete"
  )) %>%
  pivot_longer(
    cols = -EXPAGE,
    names_to = "Indicator",
    values_to = "PercentComplete",
    names_pattern = "(.*)_complete"
  ) |> 
  mutate(Data="Filled")

# Create the line plot to visualize completeness over time
ggplot(
  bind_rows(
    missingness_summary, 
    missingness_summary_filled
    ) |> 
    arrange(Indicator,EXPAGE,Data), 
  aes(
    x = EXPAGE, y = PercentComplete, 
    linetype = Data, group = Data,color = Indicator)
  ) +
  facet_wrap(.~Indicator)+
  geom_line(aes(),linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, 100), labels = scales::percent_format(scale = 1)) +
  labs(
    title = "Data Completeness of Exposure Indicators Over Time (Filled)",
    subtitle = "Percentage of non-missing observations for each indicator at each age",
    x = "Child's Age (Years)",
    y = "Percent Complete",
    color = "Indicator"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Transform Data for RMLCA

RMLCA requires data to be in a **wide format**, where each row represents one participant and columns represent the indicator variables at each time point. This section converts the data from the current long format to the required wide format.

### 1. Select and Recode Variables for Pivoting

Before pivoting, we select the key indicator variables and ensure they are in a numeric format suitable for both modeling and the pivot operation. The `mutate(across(...))` command efficiently converts all specified factor variables to their underlying numeric codes.

```{r select-and-recode}
# Select necessary columns for pivoting
# Convert factors to numeric for the modeling step
childhood_recoded_for_pivot <- childhood_data%>%
  mutate(across(c(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES), as.numeric))

glimpse(childhood_recoded_for_pivot)
```

### 2. Create Wide Dataset for Early-Life Exposure (Ages 0-5)

This chunk creates the primary analysis dataset for RMLCA. It first filters the data to include only observations where `EXPAGE` is between 0 and 5. Then, `pivot_wider()` transforms the data: `id_cols = USUBJID` ensures each participant has one row, `names_from = EXPAGE` creates new columns based on the age, and `values_from` specifies which variables' values will fill the new columns.

```{r pivot-wide-0-5}
# Filter for ages 0-5 and pivot to wide format
dt_rmlca_0_5 <- childhood_recoded_for_pivot %>%
  filter(EXPAGE >= 0 & EXPAGE <= 5) %>%
  pivot_wider(
    id_cols = USUBJID,
    names_from = EXPAGE,
    values_from = c(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4)
  )

# Inspect the final wide dataset
glimpse(dt_rmlca_0_5)
```

The resulting wide dataset for the 0-5 year period is inspected above, showing variables like `LIVEDWOM_0`, `LIVEDWOM_1`, etc.

### 3. Create Wide Dataset for Middle-Childhood Exposure (Ages 6-10)

This process is repeated for the middle-childhood period (ages 6-10) to create the dataset that will be used for the sensitivity analysis, as outlined in the SAP.

```{r pivot-wide-6-10}
# Filter for ages 6-10 and pivot to wide format
dt_rmlca_6_10 <- childhood_recoded_for_pivot %>%
  filter(EXPAGE >= 6 & EXPAGE <= 10) %>%
  pivot_wider(
    id_cols = USUBJID,
    names_from = EXPAGE,
    values_from = c(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4)
  )

# Inspect the final wide dataset
glimpse(dt_rmlca_6_10)
```

The resulting wide dataset for the 6-10 year period is inspected above.

## Descriptive Analysis of RMLCA Indicators

This section provides a summary of the indicator variables, stratified by age, using the long-format data before it was pivoted. This gives a clear year-by-year overview of the exposures.

### Table 1: Descriptive Statistics for Early-Life Indicators by Age (0-5)

The following table summarizes the distribution of each RMLCA indicator for each year from age 0 to 5. This allows us to observe trends in exposure prevalence during early childhood.

```{r describe-by-age-0-5}
# Filter for the age group and create the summary table
table_by_age_0_5 <- childhood_filtered %>%
  filter(EXPAGE >= 0 & EXPAGE <= 5) %>%
  select(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4, EXPAGE) %>%
  tbl_summary(
    by = EXPAGE,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "Missing"
  ) %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 1: Distribution of RMLCA Indicators by Age (0-5 Years)**")

table_by_age_0_5
```

### Table 2: Descriptive Statistics for Middle-Childhood Indicators by Age (6-10)

Similarly, this table summarizes the distribution of indicators for each year from age 6 to 10, providing insight into the middle-childhood exposure environment.

```{r describe-by-age-6-10}
# Filter for the age group and create the summary table
table_by_age_6_10 <- childhood_filtered %>%
  filter(EXPAGE >= 6 & EXPAGE <= 10) %>%
  select(LIVEDWOM, LIVEDWOF, EDUCMOTH, EDUCFATH, HHSES, HHCHLD4, EXPAGE) %>%
  tbl_summary(
    by = EXPAGE,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 1,
    missing_text = "Missing"
  ) %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Table 2: Distribution of RMLCA Indicators by Age (6-10 Years)**")

table_by_age_6_10
```



## Missingness Intersections (UpSet with ComplexUpset)

Now we show intersections of missingness **for each year (EXPAGE)** separately. Each UpSet plot encodes which indicators are simultaneously missing at that specific age.

```{r upset-missingness-by-year, message=FALSE, warning=FALSE, fig.height=3.5}
indicators <- c("LIVEDWOM","LIVEDWOF","EDUCMOTH","EDUCFATH","HHSES","HHCHLD4")

# Build a year-specific dataset: for each year, one row per participant with
# logical flags indicating whether each indicator is missing at that year.
miss_by_year <- function(df, y){
  df %>%
    dplyr::filter(EXPAGE == y) %>%
    dplyr::group_by(USUBJID) %>%
    dplyr::summarise(
      dplyr::across(dplyr::all_of(indicators), ~ any(is.na(.x)), .names = "{.col}"),
      .groups = "drop"
    ) %>%
    dplyr::mutate(dplyr::across(dplyr::all_of(indicators), as.logical))
}

years <- sort(unique(childhood_filtered$EXPAGE))
plots <- lapply(years, function(y){
  dat <- miss_by_year(childhood_filtered, y)
  if (nrow(dat) == 0) return(NULL)
  ComplexUpset::upset(
    dat,
    intersect = indicators,
    name = paste0("Missingness at age ", y),
    base_annotations = list(
      'Intersection size' = intersection_size(text = list(angle = 90, vjust = 0.5, hjust = 0))
    ),
    width_ratio=0.15, min_size=5, sort_intersections_by = 'degree'
  )
})

# Pair plots two-by-two using wrap_plots to avoid layout errors
plots_nonnull <- Filter(Negate(is.null), plots)
if (length(plots_nonnull) > 0) {
  for (i in seq(1, length(plots_nonnull), by = 2)) {
    p1 <- plots_nonnull[[i]]
    p2 <- if (i + 1 <= length(plots_nonnull)) plots_nonnull[[i + 1]] else NULL
    if (!is.null(p2)) {
      #print(patchwork::wrap_plots(p1, p2, ncol = 1))
      print(p1)
      print(p2)
    } else {
      print(p1)
    }
  }
}
```

```{r upset-missingness, message=FALSE, warning=FALSE, fig.height=3.5}
# If ComplexUpset is not installed, run: install.packages('ComplexUpset')

indicators <- c("LIVEDWOM","LIVEDWOF","EDUCMOTH","EDUCFATH","HHSES","HHCHLD4")

miss_any <- function(df, ages){
  df %>%
    dplyr::filter(EXPAGE %in% ages) %>%
    dplyr::group_by(USUBJID) %>%
    dplyr::summarise(
      dplyr::across(dplyr::all_of(indicators), ~ any(is.na(.x)), .names = "{.col}"),
      .groups = "drop"
    ) %>%
    dplyr::mutate(dplyr::across(dplyr::all_of(indicators), as.logical))
}

miss_0_5  <- miss_any(childhood_filtered, 0:5)
miss_6_10 <- miss_any(childhood_filtered, 6:10)

# UpSet: early-life (0–5) — use only intersection_size(text = list(angle = 90, vjust = 0.5, hjust = 0.5)) for broad compatibility
upset(
  miss_0_5,
  intersect = indicators,
  name = "Missingness (0–5 years)",
  base_annotations = list(
    'Intersection size' = intersection_size(text = list(angle = 90, vjust = 0.5, hjust = 0))
  ),
  width_ratio=0.15, min_size=0, sort_intersections_by = 'degree'
)

# UpSet: middle-childhood (6–10)
upset(
  miss_6_10,
  intersect = indicators,
  name = "Missingness (6–10 years)",
  base_annotations = list(
    'Intersection size' = intersection_size(text = list(angle = 90, vjust = 0.5, hjust = 0))
  ),
  width_ratio=0.15, min_size=0, sort_intersections_by = 'degree'
)
```

## Save RMLCA Datasets

To facilitate a reproducible workflow, we save the prepared wide-format datasets. This allows us to load them directly in the modeling script without re-running the preparation steps. The `save()` function writes the R objects to a `.RData` file in the specified output path.

```{r save-rmlca-data}

# Save the 0-5 years wide dataset
save(dt_rmlca_0_5, file = file.path(output_path, "dt_rmlca_0_5.RData"))

# Save the 6-10 years wide dataset
save(dt_rmlca_6_10, file = file.path(output_path, "dt_rmlca_6_10.RData"))
```

The two wide-format datasets, `dt_rmlcalca_0_5.RData` and `dt_rmlcalca_6_10.RData`, have been saved to the `statistical_analysis/data/adam` directory.
