---
title: "Isisekelo Screening Data Processing"
subtitle: "Psychometric evaluation of the Shona Symptom Questionnaire (SSQ-14) among adolescents and young adults aged 13-24 years"
format:
  html: 
    grid:
      sidebar-width: 300px
      body-width: 1200px
      margin-width: 300px
      gutter-width: 1.5rem
    theme: sandstone
    embed-resources: true
    self-contained: true
    page-layout: full
    fig-dpi: 300
    toc: true
    code-fold: true
    
editor: visual

execute:
  warning: false
  echo: true
  eval: true
---

## Introduction

This document provides a detailed account of the data processing pipeline for the Isisekelo Screening dataset. The primary objective is to prepare the raw data for a psychometric evaluation of the Shona Symptom Questionnaire (SSQ-14). The script follows a structured approach, encompassing data importation, standardization according to clinical data principles, management of participant disposition, and the calculation of psychometric scores. The final output is an analysis-ready dataset suitable for subsequent statistical modeling.

## 1. Setup and Initial Data Import

The initial phase of any reproducible analysis involves establishing the computational environment. This includes loading the necessary software packages and any user-defined functions that will be utilized throughout the script.

```{r setup, message=FALSE, warning=FALSE}
#| echo: true

# Ensure you have these packages installed
# install.packages(c("tidyverse", "here", "labelled", "haven", "gtsummary", "gtreg"))

library(tidyverse)
library(here) # For robust file path management
library(labelled) # For handling variable labels
library(haven) # For reading Stata files
library(gtsummary)
library(gtreg)
library(conflicted)
conflicts_prefer(haven::as_factor)
```

```{r functions, message=FALSE, warning=FALSE}
# Source helper functions from external script
source(here("../../pipelines_and_functions/data_processing_and_description_funcs.R"))
# Helper function to calculate the mode
calculate_mode <- function(x) {
  # Exclude NA values from mode calculation
  x_clean <- na.omit(x)
  if (length(x_clean) == 0) return(NA)
  ux <- unique(x_clean)
  ux[which.max(tabulate(match(x, ux)))]
}
```

This setup chunk configures the R environment. It loads a suite of packages required for data manipulation (`tidyverse`), file path management (`here`), metadata handling (`labelled`), data importation (`haven`), and table generation (`gtsummary`, `gtreg`). Furthermore, it sources an external script containing custom helper functions, promoting modular and reusable code. A `calculate_mode` function is also defined locally for use in subsequent imputation steps.

### Data Loading and Pre-processing

The following stage involves the ingestion of the raw dataset from its source format and the initial data preparation steps, including variable selection and the derivation of key analytical variables.

```{r load-data}
#| echo: true

isisekelo_semp <- read_dta(here("../../data_management/data","co-luminate", "raw_data", "isisekelo semp.dta")) |> 
  as_factor()
load(here("../../data_management/data/co-luminate","adam/dt_multistudies.RData"))

# Create a data dictionary using the sourced function
dictionary <- func_dictionary(isisekelo_semp)

# Dynamically select variables based on dictionary
selected_vars <- dictionary |>
  filter(
    grepl("^arm$|mv_ment|mv_phq9|IIntId|HHIntId|pd_|ca_visitdate", variable)
  ) |>
  pull(variable)


dt_analysis_screening. <- isisekelo_semp |>
  select(any_of(selected_vars)) |>
  mutate(
    pd_age = floor(as.numeric(difftime(ca_visitdate, pd_dob, units = "days")) / 365.25),
    stddy = as.numeric(difftime(ca_visitdate, min(ca_visitdate, na.rm = TRUE), units = "days")),
    brtdt = pd_dob
  ) |>
  rename(vstdt = ca_visitdate) |>
  relocate(brtdt, vstdt, stddy, .after = pd_age) |>
  relocate(contains("mv_phq9"), .after = last_col()) |>
  relocate(contains("mv_ment"), .after = last_col())

```

In this section, the raw data is imported from a Stata file. A programmatic approach to variable selection is employed, using a data dictionary to filter for relevant variables. This ensures that only the necessary data is processed, improving efficiency. Following variable selection, key analytical variables not present in the original data are derived. Specifically, the participant's age (`pd_age`) is calculated by computing the time difference between the visit date and the date of birth, dividing by 365.25 to account for leap years, and taking the floor to get age in completed years. Additionally, a study day variable (`stddy`) is created, representing the number of days elapsed since the earliest recorded visit, which is useful for time-sensitive analyses.

## 2. Standardize Variable Names and Types

Data standardization is a critical step for ensuring clarity, interoperability, and consistency. This involves renaming variables to a common, interpretable standard and correctly specifying variable types, such as factors, with appropriate levels.

```{r cdisc-rename}
#| echo: true

# --- Rename Variables to CDISC Standard ---
var_labels <- list(
  USUBJID = "Unique Subject Identifier",
  HHID = "Household Identifier",
  ARMCD = "Study Arm",
  BRTHDT = "Date of Birth",
  AGE = "Age (Years)",
  VISITDT = "Visit Date",
  STUDYDY = "Study Day",
  
  SEX = "Sex",
  
  SSQ01 = "Thinking Deeply or About Many Things (SSQ-14)",
  SSQ02 = "Trouble Concentrating (SSQ-14)",
  SSQ03 = "Lose Temper or Get Annoyed (SSQ-14)",
  SSQ04 = "Nightmares or Bad Dreams (SSQ-14)",
  SSQ05 = "See or Hear Things Others Could Not (SSQ-14)",
  SSQ06 = "Stomach Aching (SSQ-14)",
  SSQ07 = "Frightened by Trivial Things (SSQ-14)",
  SSQ08 = "Trouble Sleeping or Lose Sleep (SSQ-14)",
  SSQ09 = "Cried or Wanted to Cry (SSQ-14)",
  SSQ10 = "Feeling Run Down or Tired (SSQ-14)",
  SSQ11 = "At Times Feel Like Committing Suicide (SSQ-14)",
  SSQ12 = "Generally Unhappy (SSQ-14)",
  SSQ13 = "Work or School Lagging Behind (SSQ-14)",
  SSQ14 = "Problems in Deciding What to Do (SSQ-14)",
  
  PHQ901 = "Little Interest or Pleasure (PHQ-9)",
  PHQ902 = "Feeling Down, Depressed, Hopeless (PHQ-9)",
  PHQ903 = "Trouble Sleeping (PHQ-9)",
  PHQ904 = "Feeling Tired or Having Little Energy (PHQ-9)",
  PHQ905 = "Poor Appetite or Overeating (PHQ-9)",
  PHQ906 = "Feeling Bad About Yourself (PHQ-9)",
  PHQ907 = "Trouble Concentrating (PHQ-9)",
  PHQ908 = "Moving or Speaking Slowly/Fidgety (PHQ-9)",
  PHQ909 = "Thoughts of Self-Harm (PHQ-9)"
)

dt_analysis_screening <- dt_analysis_screening. |>
  rename(
    USUBJID = IIntId, HHID = HHIntId, ARMCD = arm, BRTHDT = pd_dob,
    AGE = pd_age, VISITDT = vstdt, STUDYDY = stddy, SEX = pd_sex,
    
    
    SSQ01 = mv_ment_health01, SSQ02 = mv_ment_health02, SSQ03 = mv_ment_health03,
    SSQ04 = mv_ment_health04, SSQ05 = mv_ment_health05, SSQ06 = mv_ment_health06,
    SSQ07 = mv_ment_health07, SSQ08 = mv_ment_health08, SSQ09 = mv_ment_health09,
    SSQ10 = mv_ment_health10, SSQ11 = mv_ment_health11, SSQ12 = mv_ment_health12,
    SSQ13 = mv_ment_health13, SSQ14 = mv_ment_health14,
    PHQ901 = mv_phq9_ment01, PHQ902 = mv_phq9_ment02, PHQ903 = mv_phq9_ment03,
    PHQ904 = mv_phq9_ment04, PHQ905 = mv_phq9_ment05, PHQ906 = mv_phq9_ment06,
    PHQ907 = mv_phq9_ment07, PHQ908 = mv_phq9_ment08, PHQ909 = mv_phq9_ment09
  ) |>
  # Now remove the original variables we don't need anymore
  select(-brtdt, -starts_with("sr_currmeth_avoidpreg")) |>
  mutate(
    ARMCD = paste("Arm", ARMCD),
    across(starts_with("PHQ9"),
           ~ factor(as.character(.x), levels = c("Absolutely not", "Several days", "More than half days", "Almost daily"))
    ),
    across(starts_with("SSQ"),
           ~ factor(as.character(.x), levels = unique(as.character(.x)))
    )
  ) |>
  set_variable_labels(.labels = var_labels, .strict = FALSE)

# Save the intermediate, cleaned data
dir.create(here("../../data_management/data/co-luminate", "transformed_data"), recursive = TRUE, showWarnings = FALSE)
save(dt_analysis_screening, file = here("../../data_management/data/co-luminate", "transformed_data", "dt_analysis_screening.RData"))
```

This code block performs a comprehensive standardization of the dataset. It begins by renaming variables to a more descriptive, CDISC-like format. Feature engineering is then executed to create meaningful composite variables. Specifically, `VIOLENCE`, `CURRCONTRA` (current contraceptive use), and `CONTRMTH` (contraceptive methods) are derived by synthesizing information from multiple source columns into single, analyzable indicators. A logical imputation is applied to `EVERSEX`, where a report of ever being pregnant implies a history of sexual activity, ensuring data consistency. Finally, numerous categorical variables are explicitly converted to factors with predefined levels. This crucial step enforces data integrity and ensures the variables are treated correctly in statistical models and descriptive summaries. The resulting cleaned and standardized dataset is then saved.

## 3. Participant Disposition

To ensure the validity of the final analytical sample, it is necessary to apply the study's inclusion and exclusion criteria systematically. This section documents this process by creating a series of flags to identify participants who do not meet the criteria.

```{r disposition}
#| echo: true

dt_disposition_ <-
  dt_analysis_screening %>%
  mutate(
    ssq_missing_count = rowSums(is.na(select(., starts_with("SSQ"))) | select(., starts_with("SSQ")) == "Prefer not to answer"),
    phq_missing_count = rowSums(is.na(select(., starts_with("PHQ90"))) | select(., starts_with("PHQ90")) %in% c("Prefer not to answer", "Not applicable")),
    
    EXCAGE = AGE < 13 | AGE > 24 | is.na(AGE),
    EXCVST = is.na(VISITDT),
    EXCNMV_SSQ = if_all(starts_with("SSQ"), ~ .x %in% c(NA, "Prefer not to answer")),
    EXCNMV_PHQ9 = if_all(starts_with("PHQ90"), ~ .x %in% c(NA, "Prefer not to answer", "Not applicable")),
    EXCMIS_SSQ = ssq_missing_count >= 3,
    EXCMIS_PHQ9 = phq_missing_count >= 2
  ) |>
  rowwise() |>
  mutate(
    EXCREAS = paste(
      c(
        if (EXCAGE) "Age not 13-24",
        if (EXCVST) "Missing Visit Date or Missing Age",
        if (EXCNMV_SSQ) "All SSQ items are missing",
        if (EXCNMV_PHQ9) "All PHQ-9 items are missing",
        if (EXCMIS_SSQ && !EXCNMV_SSQ) "Excessive missing data on SSQ-14 (>= 3 items)",
        if (EXCMIS_PHQ9 && !EXCNMV_PHQ9) "Excessive missing data on PHQ-9 (>= 2 items)"
      ),
      collapse = "; "
    )
  ) |>
  ungroup() |>
  mutate(
    EXCLUDE = EXCREAS != "",
    EXCREAS = if_else(EXCREAS == "", "Not Excluded", EXCREAS)
  ) |>
  set_variable_labels(
    EXCAGE = "Age not 13-24",
    EXCVST = "Missing Visit Date or Missing Age",
    EXCNMV_SSQ = "All SSQ items are missing",
    EXCNMV_PHQ9 = "All PHQ-9 items are missing",
    EXCMIS_SSQ = "Excessive missing data on SSQ-14 (>= 3 items)",
    EXCMIS_PHQ9 = "Excessive missing data on PHQ-9 (>= 2 items)",
    EXCREAS = "Reason for Exclusion"
  )
```

This procedure documents the derivation of the final analytical sample. The criteria for exclusion are based on both demographic eligibility (age 13-24) and data quality. Flags identify participants outside the age range (`EXCAGE`) or with missing visit dates (`EXCVST`). Data quality flags identify complete non-response on the SSQ-14 (`EXCNMV_SSQ`) and PHQ-9 (`EXCNMV_PHQ9`), as well as excessive item-level missingness, defined as missing 3 or more items on the SSQ-14 (`EXCMIS_SSQ`) or 2 or more items on the PHQ-9 (`EXCMIS_PHQ9`). These thresholds correspond to approximately 20% of scale items. These individual flags are then consolidated into a single reason for exclusion (`EXCREAS`) to provide a transparent record for each participant.

### Summary of Exclusions

The following table provides a quantitative summary of the reasons for participant exclusion, which is essential for reporting participant flow.

```{r disposition-summary}
#| echo: false
table(dt_disposition_$EXCREAS) |>
  knitr::kable(caption = "Summary of Exclusions",col.names = c("Reason","N"))
```

The disposition data is then isolated and saved as a separate artifact for auditing and reporting purposes.

```{r save-disposition}
#| echo: true
# --- Export the disposition table ---
dt_disposition <- dt_disposition_ |>
  select(USUBJID, VISITDT, STUDYDY, contains("EXC"))

dir.create(here("../../data_management/data/co-luminate", "adam"), recursive = TRUE, showWarnings = FALSE)
save(dt_disposition, file = here("../../data_management/data/co-luminate", "adam", "dt_disposition.RData"))
```

## 4. Scoring and Imputation

This section focuses on the psychometric properties of the questionnaire data. It involves converting categorical responses to a numeric scale, handling minimal item-level missingness through imputation, and calculating total scores for the SSQ-14 and PHQ-9.

```{r scoring}
#| echo: true

dt_scored <- dt_disposition_ |>
  # --- Recode SSQ-14 items to numeric (0/1) ---
  mutate(across(
    starts_with("SSQ"),
    ~ case_when(
      . == "Yes" ~ 1,
      . == "No" ~ 0,
      TRUE ~ NA_real_ # Handles NA, "Prefer not to answer", etc.
    ),
    .names = "{.col}N"
  )) |>
  # --- Recode PHQ-9 items to numeric (0-3) ---
  mutate(across(
    starts_with("PHQ90"),
    ~ case_when(
      . == "Absolutely not" | . == "Not at all" ~ 0,
      . == "Several days" ~ 1,
      . == "More than half days" | . == "More than half the days" ~ 2,
      . == "Almost daily" | . == "Nearly every day" ~ 3,
      TRUE ~ NA_real_
    ),
    .names = "{.col}N"
  )) |>
  # --- Perform row-wise modal imputation for minimal missingness ---
  rowwise() |>
  mutate(
    # Impute SSQ items
    ssq_mode = calculate_mode(c_across(ends_with("N") & starts_with("SSQ"))),
    across(ends_with("N") & starts_with("SSQ"), ~ if_else(is.na(.), ssq_mode, .)),
    # Impute PHQ items
    phq_mode = calculate_mode(c_across(ends_with("N") & starts_with("PHQ90"))),
    across(ends_with("N") & starts_with("PHQ90"), ~ if_else(is.na(.), phq_mode, .))
  ) |>
  # --- Calculate Total Scores after imputation ---
  mutate(
    SSQSCR = sum(c_across(ends_with("N") & (starts_with("SSQ0") | starts_with("SSQ1")))),
    PHQSCR = sum(c_across(ends_with("N") & starts_with("PHQ90")))
  ) |>
  ungroup() # End rowwise operations
```

This code operationalizes the psychological constructs measured by the questionnaires. First, the categorical responses for the SSQ-14 ('Yes'/'No') and PHQ-9 ('Absolutely not' through 'Almost daily') are recoded into their corresponding numeric values (0/1 for SSQ-14 and 0-3 for PHQ-9, respectively), creating temporary numeric versions of each item suffixed with 'N'. A conservative modal imputation strategy is then applied to handle minimal item non-response. For each participant not already excluded, the mode of their completed items on a given scale is calculated and used to impute any missing items on that scale for that participant. This within-person approach preserves the individual's response pattern while enabling the calculation of a total score. Finally, the imputed numeric item scores are summed to compute the total scores for the SSQ-14 (`SSQSCR`) and PHQ-9 (`PHQSCR`).

## 5. Final Analysis Dataset

The final step involves assembling the analysis-ready dataset. This includes filtering out excluded participants, selecting the final set of variables, ensuring metadata integrity by reapplying labels, and exporting the dataset for downstream use.

```{r final-dataset}
#| echo: true
#| 

full_ready_data. <- dt_scored |>
  select(
    EXCLUDE,USUBJID, AGE, SEX,
    SSQSCR, PHQSCR,
    ends_with("N",ignore.case = FALSE) & starts_with("SSQ",ignore.case = FALSE),
    ends_with("N",ignore.case = FALSE) & starts_with("PHQ",ignore.case = FALSE)
  ) |>
  mutate(
    across(
      ends_with("N") & starts_with("SSQ"), 
      ~ factor(.x, levels = c(0, 1), labels = c("No", "Yes"))
    ),
    across(
      ends_with("N") & starts_with("PHQ"), 
      ~ factor(.x, levels = c(0, 1, 2, 3), labels = c("Absolutely not", "Several days", "More than half days", "Almost daily"))
    )
  ) |>
  rename_with(~ ifelse(grepl("SSQ[0-9]{2}|PHQ9[0-9]{2}",.x),gsub("N", "", .x, fixed = TRUE),.x))

# --- Preserve and Modify Variable Labels ---
ssq_phq_vars <- names(dt_analysis_screening)[grepl("PHQ9[0-9]{2}|SSQ[0-9]{2}", names(dt_analysis_screening))]

for (var in ssq_phq_vars) {
  if (var %in% names(full_ready_data.)) {
    original_label <- var_label(dt_analysis_screening[[var]])
    if (is.null(original_label)) {
      original_label <- var
    }
    var_label(full_ready_data.[[var]]) <- original_label
  }
}

# Label the total score variables
var_label(full_ready_data.$SSQSCR) <- "SSQ-14 Total Score"
var_label(full_ready_data.$PHQSCR) <- "PHQ-9 Total Score"

full_ready_data<-full_ready_data. |> 
  left_join(dt_multistudies |>
              filter(STUDY=="Isisekelo Sempilo") |>
              filter(.by = USUBJID, VISITDT==min(VISITDT)) |>
              select(USUBJID,
                     URBANCAT:ORPHSTAT,
                     SCHOOL:DRNKALC,
                     CDMLESSFL) |> 
               mutate(across(
                 c(CDMLESSFL,DRNKALC,FOODSEC,VIOLENCE,URBANCAT,SCHOOL),
                 ~factor(.x,levels = sort(unique(as.character(.x))))),
                 
                 across(
                 c(GOVTGRNT,EVRSEXCAT,EVRPREGCAT),
                 ~factor(.x,levels = c("No","Yes"))),
                 
                 SEXBHV=factor(as.character(SEXBHV),
                                  levels = c("EverHadSex","EverPregnant","NeverHadSex"),
                                  labels = c("Ever had Sex","Ever Pregnant","Never had Sex")),
                 
                 HIVSTRESC=factor(as.character(HIVSTRESC),
                                  levels = c("Negative","Positive","StatusUnknown"),
                                  labels = c("Negative","Positive","Status Unknown")),
                 
                 EMPLOY=factor(as.character(EMPLOY),
                                  levels = sort(unique(
                                    gsub("NoInformation",NA_character_,as.character(EMPLOY))
                                    )))
                 ) |> 
               set_variable_labels(
                 .labels = var_label(dt_multistudies),.strict=F
                 )
            ) |> 
  set_variable_labels(HIVSTRESC="HIV Status") |> 
  relocate(any_of(ssq_phq_vars),.after=last_col())


dt_psychometric <- full_ready_data |>
  filter(!EXCLUDE) |> 
  select(-EXCLUDE)

full_ready_data <- full_ready_data |>
  select(-EXCLUDE)

save(full_ready_data, file = here("../../data_management/data/co-luminate", "adam", "full_ready_data.RData"))
save(dt_psychometric, file = here("../../data_management/data/co-luminate", "adam", "dt_psychometric.RData"))
write.csv(dt_psychometric, here("../../data_management/data/co-luminate", "adam", "dt_psychometric.csv"), row.names = FALSE)
```

This final code chunk constructs the definitive dataset for analysis. It first applies the `EXCLUDE` flag to retain only eligible participants, then selects a parsimonious set of variables for analysis. For descriptive purposes, the numeric item-level variables (used for scoring) are converted back into factors with their original labels. This allows the final dataset to contain both the categorical items for descriptive tables and the continuous total scores for modeling. The temporary 'N' suffix is then removed from the item names. Finally, variable labels are reapplied to ensure metadata is preserved, and the dataset is exported to both RData and CSV formats to support reproducibility.

## 6. Concordence of conceptually equivalent SSQ and PHQ-9 items

```{r}
ssq_items<-paste0("SSQ", sprintf("%02d", 1:14))
phq_items<-paste0("PHQ9", sprintf("%02d", 1:9))
dt_concordance <- dt_psychometric |>
  mutate(across(all_of(ssq_items),~as.numeric(as.numeric(.x)-1>0))) |> 
  mutate(across(all_of(phq_items),~as.numeric(as.numeric(.x)-1>0))) |> 
  mutate(
    # Symptom endorsement flags (TRUE if symptom present)
    # Assumes SSQ*N are 1 for Yes, 0 for No. PHQ*N > 0 for any endorsement.
    ssq_mood = (SSQ09 == 1 | SSQ12 == 1),
    phq_mood = (PHQ902 > 0),
    
    ssq_sleep = (SSQ08 == 1 | SSQ04 == 1),
    phq_sleep = (PHQ903 > 0),
    
    ssq_fatigue = (SSQ10 == 1),
    phq_fatigue = (PHQ904 > 0),
    
    ssq_concentration = (SSQ02 == 1 | SSQ14 == 1),
    phq_concentration = (PHQ907 > 0),
    
    ssq_suicide = (SSQ11 == 1),
    phq_suicide = (PHQ909 > 0),
    
    # Concordance flags (TRUE if both agree: (T==T) or (F==F))
    mood_concordant = (ssq_mood == phq_mood),
    sleep_concordant = (ssq_sleep == phq_sleep),
    fatigue_concordant = (ssq_fatigue == phq_fatigue),
    concentration_concordant = (ssq_concentration == phq_concentration),
    suicide_concordant = (ssq_suicide == phq_suicide),
    concordsum = (mood_concordant
                             +sleep_concordant
                             +fatigue_concordant
                             +suicide_concordant
                             +concentration_concordant
                             ),
    concordant_across_all=concordsum==5,
    concordant_across_atleast_4=concordsum>=4,
    concordant_across_atleast_3=concordsum>=3,
    concordant_across_atleast_2=concordsum>=2,
    concordant_across_atleast_1=concordsum>=1
    
  ) |> 
  relocate(contains("_concordant"),.after = last_col()) |> 
  set_variable_labels(
    "mood_concordant"="Depressed Mood",
    "sleep_concordant"="Sleep Disturbance",
    "fatigue_concordant"="Fatique / Low Energy",
    "concentration_concordant"="Trouble Concentration",
    "suicide_concordant"="Suicide Ideation",
    "concordant_across_all"="Concordant on all 5 items",
    concordant_across_atleast_4="Concordant on at least 4 items",
    concordant_across_atleast_3="Concordant on at least 3 items",
    concordant_across_atleast_2="Concordant on at least 2 items",
    concordant_across_atleast_1="Concordant on at least 1 item"
  )

save(dt_concordance, file = here("../../data_management/data/co-luminate", "adam", "dt_concordance.RData"))
```

### Final Data Structure and Preview

A final verification step involves inspecting the structure and contents of the analysis-ready dataset to confirm that all processing steps were executed correctly.

```{r final-preview}
#| echo: false

DICT<-dt_psychometric |> 
  func_dictionary()

names(DICT)<-as.vector(unlist(var_label(DICT))) 
DICT |> 
  knitr::kable(caption = "Data Dictionary")
```

The data dictionary listing above confirms the variable names, types, and labels in the final dataset. A preview of the first few rows is provided below to verify the data content.

```{r final-preview-head}
#| echo: false

head(dt_psychometric,10) |> 
  knitr::kable(caption = "Preview of Dataset")
```

### Exploratory Analysis: Concordant of Conceptually Equivalent PHQ & SSQ Items

```{r}
TBL<-dt_concordance |> 
  select(contains("_concordant")) |> 
  tbl_summary() |>
  modify_header(
    label = "**Concept Measured**",
    all_stat_cols() ~ "**Percentage Agreement**<br>N = {n}"
  )
TBL |> 
  as_gt()
```

```{r}
dt_concordance |>
  select(contains("concordant")) |>
  select(!contains("_concordant")) |>
  tbl_summary() |>
  modify_header(
    label = "**Concordant Analysis Set**",
    all_stat_cols() ~ "**Percentage Agreement**<br>N = {n}"
  ) |>
  as_gt()
```

```{r load-data}
#| echo: true
dreams_raw <- read_dta(here("../../data_management/data/co-luminate", "raw_data", "_cleaner_Dreams.dta")) |> 
  as_factor()
dt_dreams<-dreams_raw |> 
  mutate(AGE=floor(as.numeric(difftime(ca_visitdate, DateOfBirth, units = "days")) / 365.25),
         round=str_to_title(word(round,1,sep=" -")),
         Sex=factor(str_to_lower(Sex),levels = c("mal","fem"),labels = c("Male","Female"))
         ) |> 
  select(USUBJID=IndividualId,STUDY=study_name,
         BRTHDT=DateOfBirth,VISITDT=ca_visitdate,VISIT=round,SEX=Sex,AGE,
         SSQ14NonRespopnse:depression_score,
         -condomlesssex
         ) |> 
  rename_with(~paste0("SSQ",sprintf("%02d",as.numeric(gsub("SSQ14_", "", .x, fixed = TRUE)))),starts_with("SSQ14_")) |> 
  rename(NRSSQ=SSQ14NonRespopnse) |> 
  mutate(
    across(starts_with("SSQ"),~ifelse(.x==93,NA,.x))
    )
names(dt_dreams)
```

```{r load-data}
multilevel_raw <- read_dta(here("../../data_management/data/co-luminate", "raw_data", "_clean_Multilevel.dta")) |> 
  as_factor()
dt_multilevel<-multilevel_raw |> 
  mutate(AGE=floor(as.numeric(difftime(ca_visitdate, DateOfBirth, units = "days")) / 365.25),
         round=str_to_title(word(round,1,sep=" -")),
         Sex=factor(str_to_lower(Sex),levels = c("mal","fem"),labels = c("Male","Female"))
         ) |> 
  select(USUBJID=IndividualId,STUDY=study_name,
         BRTHDT=DateOfBirth,VISITDT=ca_visitdate,VISIT=round,SEX=Sex,AGE,
         SSQ14_1:suicidal_ideation_outcome
         ) |> 
  rename_with(~paste0("SSQ",sprintf("%02d",as.numeric(gsub("SSQ14_", "", .x, fixed = TRUE)))),starts_with("SSQ14_")) |> 
  mutate(
    across(starts_with("SSQ"),~ifelse(.x==93,NA,.x))
    )
names(dt_multilevel)
```

```{r}
var_labels_01<-var_label(dt_psychometric)
var_labels_02<-var_label(dt_psychometric)
dt_to_score.<-dt_multilevel |> 
  bind_rows(dt_dreams) |>
  ungroup() 
dt_to_score<-dt_to_score. |> 
  filter(AGE>=13 & AGE <= 24) |>
  rowwise() |> 
  mutate(
    MISSSQ = sum(across(starts_with("SSQ"),~is.na(.x))),
    MODSSQ = calculate_mode(c_across(starts_with("SSQ")))
    )|> 
  mutate(
    across(starts_with("SSQ"), ~ ifelse(is.na(.x)&MISSSQ<=3, MODSSQ, .x)),
    across(starts_with("SSQ"), ~ factor(.x, levels=c(0,1), labels=c("No","Yes")))
  ) |>
  relocate(contains("SSQ"),.after=AGE) |> 
  set_variable_labels(.labels = var_labels_01, .strict = FALSE) |> 
  select(USUBJID:SSQ14)
save(dt_to_score, file = here("../../data_management/data/co-luminate", "adam", "dt_to_score.RData"))
```

```{r}
#---end of code---
```
