---
title: "Dashboard"
format: 
  dashboard
---

```{r}
#| include: false
#| warning: true
#| message: false
has_dt <- requireNamespace("DT", quietly = TRUE)
has_crosstalk <- requireNamespace("crosstalk", quietly = TRUE)
has_htmltools <- requireNamespace("htmltools", quietly = TRUE)
has_dygraphs <- requireNamespace("dygraphs", quietly = TRUE)
has_xts <- requireNamespace("xts", quietly = TRUE)
has_interactive_plots <- has_dygraphs && has_xts

if (!has_htmltools) {
  stop("Package 'htmltools' is required to render dashboard.qmd.")
}
if (!has_dt) {
  warning("Package 'DT' is not installed. Rendering static tables.")
}
if (!has_crosstalk) {
  warning("Package 'crosstalk' is not installed. Rendering non-interactive filter panels.")
}
if (!has_interactive_plots) {
  warning("Packages 'dygraphs' and/or 'xts' are not installed. Interactive plots will be skipped.")
}

read_obj <- function(filename) {
  candidates <- c(
    file.path("_tools", "_objects", filename),
    file.path("..", "_tools", "_objects", filename)
  )
  idx <- which(file.exists(candidates))[1]
  if (is.na(idx)) {
    stop("Missing object file: ", filename)
  }
  readRDS(candidates[idx])
}

kpi <- read_obj("dashboard_kpi.rds")
study_summary <- read_obj("dashboard_study_summary.rds")
demog_summary <- read_obj("dashboard_demog_summary.rds")
year_summary <- read_obj("dashboard_year_summary.rds")
mediator_summary <- read_obj("dashboard_mediator_summary.rds")
filter_schema <- read_obj("dashboard_filter_schema.rds")
filter_cube <- read_obj("dashboard_filter_cube.rds")
trajectory_trend_summary <- read_obj("dashboard_trajectory_trend_summary.rds")
trajectory_pairwise_summary <- read_obj("dashboard_trajectory_pairwise_summary.rds")
dashboard_build_info <- read_obj("dashboard_build_info.rds")

fmt_int <- function(x) format(round(x), big.mark = ",", trim = TRUE)
fmt_pct <- function(x) ifelse(is.na(x), "NA", sprintf("%.1f%%", x))

date_index_from_int <- function(x, base_year = 2000L) {
  as.Date(sprintf("%d-01-01", base_year + as.integer(x)))
}

date_index_seq <- function(n, base_year = 2000L) {
  as.Date(sprintf("%d-01-01", base_year + seq_len(n) - 1L))
}

datatable_safe <- function(data, ...) {
  if (has_dt) {
    return(DT::datatable(data, ...))
  }

  if (has_crosstalk && inherits(data, "SharedData")) {
    data <- data$origData()
  }

  knitr::kable(as.data.frame(data), row.names = FALSE)
}

card <- function(label, value, note = "") {
  htmltools::div(
    class = "dashboard-kpi-card",
    htmltools::tags$p(class = "dashboard-kpi-label", label),
    htmltools::tags$p(class = "dashboard-kpi-value", value),
    htmltools::tags$p(class = "dashboard-kpi-note", note)
  )
}
```

## Dashboard Views {.tabset}

### All

This dashboard follows a standard structure: filter panel, KPI tiles, interactive plots, then detailed aggregated tables.

All outputs are built from aggregated objects only. No participant-level rows are stored in this website project.

#### Global Filter Panel

```{r}
#| echo: false
available_schema <- filter_schema[filter_schema$available, , drop = FALSE]

if (nrow(filter_cube) == 0) {
  cat("No rows available in the suppressed filter cube.")
} else {
  display_cols <- c(
    "study", "sex", "age_band",
    available_schema$column_name[!is.na(available_schema$column_name)],
    "n", "depressed_n", "prevalence_pct"
  )
  display_cols <- unique(display_cols[display_cols %in% names(filter_cube)])
  cube_display <- filter_cube[, display_cols, drop = FALSE]

  if (has_crosstalk) {
    cube_shared <- crosstalk::SharedData$new(
      cube_display,
      key = as.character(seq_len(nrow(cube_display))),
      group = "all_cube_group"
    )

    controls <- list(
      crosstalk::filter_select("all_cube_study", "Study", cube_shared, ~study, multiple = TRUE),
      crosstalk::filter_select("all_cube_sex", "Sex", cube_shared, ~sex, multiple = TRUE),
      crosstalk::filter_select("all_cube_age", "Age Band", cube_shared, ~age_band, multiple = TRUE)
    )

    if (nrow(available_schema) > 0) {
      for (i in seq_len(nrow(available_schema))) {
        col_nm <- available_schema$column_name[[i]]
        if (!is.na(col_nm) && col_nm %in% names(cube_display)) {
          controls[[length(controls) + 1]] <- crosstalk::filter_select(
            paste0("all_cube_", col_nm),
            paste0(available_schema$domain[[i]], ": ", available_schema$label[[i]]),
            cube_shared,
            stats::as.formula(paste0("~", col_nm)),
            multiple = TRUE
          )
        }
      }
    }

    do.call(
      htmltools::div,
      c(list(class = "dashboard-filter-grid"), controls)
    )

    datatable_safe(
      cube_shared,
      rownames = FALSE,
      options = list(pageLength = 8, autoWidth = TRUE, scrollX = TRUE)
    )
  } else {
    datatable_safe(
      cube_display,
      rownames = FALSE,
      options = list(pageLength = 8, autoWidth = TRUE, scrollX = TRUE)
    )
  }
}
```

#### Snapshot KPIs

```{r}
#| echo: false
htmltools::div(
  class = "dashboard-kpi-grid",
  card("Participants (snapshot)", fmt_int(kpi$participants_total), "One record per participant."),
  card("Records (analysis rows)", fmt_int(kpi$records_total), "Rows with non-missing DPBN."),
  card("Cohorts represented", fmt_int(kpi$studies_total), "Distinct studies."),
  card("Depression prevalence", fmt_pct(kpi$depression_prev_pct), "Participant snapshot prevalence."),
  card("Filter dimensions", fmt_int(kpi$filter_dimensions_available), "Available SDoH/trajectory filters."),
  card("Trajectory dimensions", fmt_int(kpi$trajectory_dimensions_available), "Configured trajectory dimensions available.")
)
```

#### Depression Prevalence Trend by Visit Year

```{r}
#| echo: false
if (!has_interactive_plots) {
  cat("Interactive plot dependencies are unavailable in this render environment.")
} else if (nrow(year_summary) == 0) {
  cat("No visit-year trend rows available after suppression.")
} else {
  clyr_vals <- sort(unique(year_summary$clyr))
  study_vals <- sort(unique(year_summary$study))

  trend_mat <- matrix(
    NA_real_,
    nrow = length(clyr_vals),
    ncol = length(study_vals),
    dimnames = list(as.character(clyr_vals), study_vals)
  )

  for (i in seq_len(nrow(year_summary))) {
    rr <- year_summary[i, , drop = FALSE]
    trend_mat[as.character(rr$clyr), as.character(rr$study)] <- as.numeric(rr$prevalence_pct)
  }

  trend_xts <- xts::xts(trend_mat, order.by = date_index_from_int(clyr_vals))

  dygraphs::dygraph(
    trend_xts,
    main = "Depression Prevalence (%) by Visit Year and Cohort"
  ) |>
    dygraphs::dyOptions(drawPoints = TRUE, pointSize = 2.8, strokeWidth = 2) |>
    dygraphs::dyRangeSelector() |>
    dygraphs::dyLegend(show = "follow") |>
    dygraphs::dyAxis("y", label = "Depression prevalence (%)") |>
    dygraphs::dyAxis("x", label = "Visit year")
}
```

#### Age-Band Depression Profile by Sex

```{r}
#| echo: false
if (!has_interactive_plots) {
  cat("Interactive plot dependencies are unavailable in this render environment.")
} else if (nrow(demog_summary) == 0) {
  cat("No demographic summary rows available after suppression.")
} else {
  age_levels <- c("13-15", "16-18", "19-21", "22-24", "25+", "Unknown")
  age_vals <- age_levels[age_levels %in% unique(as.character(demog_summary$age_band))]
  sex_vals <- sort(unique(as.character(demog_summary$sex)))

  demog_agg <- stats::aggregate(
    cbind(n, depressed_n) ~ age_band + sex,
    data = demog_summary,
    FUN = sum
  )
  demog_agg$prevalence_pct <- with(demog_agg, ifelse(n > 0, round(100 * depressed_n / n, 1), NA_real_))

  demo_mat <- matrix(
    NA_real_,
    nrow = length(age_vals),
    ncol = length(sex_vals),
    dimnames = list(age_vals, sex_vals)
  )

  for (i in seq_len(nrow(demog_agg))) {
    rr <- demog_agg[i, , drop = FALSE]
    age_key <- as.character(rr$age_band)
    sex_key <- as.character(rr$sex)
    if (age_key %in% rownames(demo_mat) && sex_key %in% colnames(demo_mat)) {
      demo_mat[age_key, sex_key] <- as.numeric(rr$prevalence_pct)
    }
  }

  demo_xts <- xts::xts(demo_mat, order.by = date_index_seq(length(age_vals)))

  dygraphs::dygraph(
    demo_xts,
    main = "Depression Prevalence (%) by Age Band and Sex"
  ) |>
    dygraphs::dyOptions(drawPoints = TRUE, pointSize = 3, strokeWidth = 2) |>
    dygraphs::dyLegend(show = "follow") |>
    dygraphs::dyAxis("y", label = "Depression prevalence (%)") |>
    dygraphs::dyAxis("x", label = "Age-band index")
}
```

#### Cohort Snapshot Table

```{r}
#| echo: false
study_tbl <- study_summary
study_tbl$participants <- fmt_int(study_tbl$participants)
study_tbl$depressed_n <- fmt_int(study_tbl$depressed_n)
study_tbl$prevalence_pct <- fmt_pct(study_tbl$prevalence_pct)
study_tbl$female_pct <- fmt_pct(study_tbl$female_pct)

datatable_safe(
  study_tbl,
  rownames = FALSE,
  filter = "top",
  colnames = c(
    "Study", "Participants", "Depressed N", "Depression Prev",
    "Female Share", "Median Age", "Age Q1", "Age Q3"
  ),
  options = list(pageLength = 8, autoWidth = TRUE, scrollX = TRUE)
)
```

### SDoH

#### Mediator Prevalence Profile by Cohort

```{r}
#| echo: false
if (!has_interactive_plots) {
  cat("Interactive plot dependencies are unavailable in this render environment.")
} else if (nrow(mediator_summary) == 0) {
  cat("No mediator summary rows available after suppression.")
} else {
  study_vals <- sort(unique(mediator_summary$study))
  med_vals <- sort(unique(mediator_summary$mediator))

  med_mat <- matrix(
    NA_real_,
    nrow = length(study_vals),
    ncol = length(med_vals),
    dimnames = list(study_vals, med_vals)
  )

  for (i in seq_len(nrow(mediator_summary))) {
    rr <- mediator_summary[i, , drop = FALSE]
    med_mat[as.character(rr$study), as.character(rr$mediator)] <- as.numeric(rr$yes_pct)
  }

  med_xts <- xts::xts(med_mat, order.by = date_index_seq(length(study_vals)))

  dygraphs::dygraph(
    med_xts,
    main = "Mediator 'Yes' Prevalence (%) by Cohort"
  ) |>
    dygraphs::dyOptions(drawPoints = TRUE, pointSize = 2.5, strokeWidth = 1.8) |>
    dygraphs::dyLegend(show = "follow") |>
    dygraphs::dyAxis("y", label = "Mediator prevalence (%)") |>
    dygraphs::dyAxis("x", label = "Cohort index")
}
```

#### Interactive Mediator Table

```{r}
#| echo: false
if (has_crosstalk) {
  med_shared <- crosstalk::SharedData$new(
    mediator_summary,
    key = ~paste(study, mediator),
    group = "med_group"
  )

  crosstalk::bscols(
    widths = c(2, 2, 8),
    crosstalk::filter_select("med_study", "Study", med_shared, ~study, multiple = TRUE),
    crosstalk::filter_select("med_name", "Mediator", med_shared, ~mediator, multiple = TRUE),
    datatable_safe(
      med_shared,
      rownames = FALSE,
      options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
      colnames = c("Study", "Mediator", "Complete N", "Yes N", "Yes (%)")
    )
  )
} else {
  datatable_safe(
    mediator_summary,
    rownames = FALSE,
    options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
    colnames = c("Study", "Mediator", "Complete N", "Yes N", "Yes (%)")
  )
}
```

#### SDoH Multi-filter Cube

```{r}
#| echo: false
available_schema <- filter_schema[filter_schema$available, , drop = FALSE]
sdoh_schema <- available_schema[tolower(available_schema$domain) == "sdoh", , drop = FALSE]

if (nrow(filter_cube) == 0) {
  cat("No rows available in the suppressed filter cube.")
} else {
  display_cols <- c(
    "study", "sex", "age_band",
    sdoh_schema$column_name[!is.na(sdoh_schema$column_name)],
    "n", "depressed_n", "prevalence_pct"
  )
  display_cols <- unique(display_cols[display_cols %in% names(filter_cube)])
  cube_display <- filter_cube[, display_cols, drop = FALSE]

  if (has_crosstalk) {
    cube_shared <- crosstalk::SharedData$new(
      cube_display,
      key = as.character(seq_len(nrow(cube_display))),
      group = "sdoh_cube_group"
    )

    controls <- list(
      crosstalk::filter_select("sdoh_cube_study", "Study", cube_shared, ~study, multiple = TRUE),
      crosstalk::filter_select("sdoh_cube_sex", "Sex", cube_shared, ~sex, multiple = TRUE),
      crosstalk::filter_select("sdoh_cube_age", "Age Band", cube_shared, ~age_band, multiple = TRUE)
    )

    if (nrow(sdoh_schema) > 0) {
      for (i in seq_len(nrow(sdoh_schema))) {
        col_nm <- sdoh_schema$column_name[[i]]
        if (!is.na(col_nm) && col_nm %in% names(cube_display)) {
          controls[[length(controls) + 1]] <- crosstalk::filter_select(
            paste0("sdoh_cube_", col_nm),
            sdoh_schema$label[[i]],
            cube_shared,
            stats::as.formula(paste0("~", col_nm)),
            multiple = TRUE
          )
        }
      }
    }

    do.call(
      htmltools::div,
      c(list(class = "dashboard-filter-grid"), controls)
    )

    datatable_safe(
      cube_shared,
      rownames = FALSE,
      options = list(pageLength = 12, autoWidth = TRUE, scrollX = TRUE)
    )
  } else {
    datatable_safe(
      cube_display,
      rownames = FALSE,
      options = list(pageLength = 12, autoWidth = TRUE, scrollX = TRUE)
    )
  }
}
```

### Trajectories

#### Trajectory Summary

```{r}
#| echo: false
if (nrow(trajectory_trend_summary) == 0) {
  cat("No trajectory trend rows available after aggregation and suppression.")
} else {
  traj_summary <- stats::aggregate(
    cbind(n, depressed_n) ~ trajectory_label,
    data = trajectory_trend_summary,
    FUN = sum
  )
  traj_summary$prevalence_pct <- with(
    traj_summary,
    ifelse(n > 0, round(100 * depressed_n / n, 1), NA_real_)
  )
  names(traj_summary) <- c("Trajectory Dimension", "N", "Depressed N", "Depression Prev (%)")
  datatable_safe(traj_summary, rownames = FALSE, options = list(pageLength = 10, autoWidth = TRUE))
}
```

#### Trajectory Trend Chart

```{r}
#| echo: false
if (!has_interactive_plots) {
  cat("Interactive plot dependencies are unavailable in this render environment.")
} else if (nrow(trajectory_trend_summary) == 0) {
  cat("No trajectory trend rows available after aggregation and suppression.")
} else {
  trend_agg <- stats::aggregate(
    cbind(n, depressed_n) ~ trajectory_label + clyr + trajectory_class,
    data = trajectory_trend_summary,
    FUN = sum
  )
  trend_agg$prevalence_pct <- with(
    trend_agg,
    ifelse(n > 0, round(100 * depressed_n / n, 1), NA_real_)
  )

  primary_label <- sort(unique(as.character(trend_agg$trajectory_label)))[1]
  plot_df <- trend_agg[as.character(trend_agg$trajectory_label) == primary_label, , drop = FALSE]

  clyr_vals <- sort(unique(plot_df$clyr))
  class_vals <- sort(unique(as.character(plot_df$trajectory_class)))

  trend_mat <- matrix(
    NA_real_,
    nrow = length(clyr_vals),
    ncol = length(class_vals),
    dimnames = list(as.character(clyr_vals), class_vals)
  )

  for (i in seq_len(nrow(plot_df))) {
    rr <- plot_df[i, , drop = FALSE]
    trend_mat[as.character(rr$clyr), as.character(rr$trajectory_class)] <- as.numeric(rr$prevalence_pct)
  }

  trend_xts <- xts::xts(trend_mat, order.by = date_index_from_int(clyr_vals))

  dygraphs::dygraph(
    trend_xts,
    main = paste0("Trajectory Trend: ", primary_label)
  ) |>
    dygraphs::dyOptions(drawPoints = TRUE, pointSize = 2.8, strokeWidth = 2) |>
    dygraphs::dyRangeSelector() |>
    dygraphs::dyLegend(show = "follow") |>
    dygraphs::dyAxis("y", label = "Depression prevalence (%)") |>
    dygraphs::dyAxis("x", label = "Visit year")
}
```

#### Interactive Trajectory Tables

```{r}
#| echo: false
if (nrow(trajectory_trend_summary) == 0) {
  cat("No trajectory trend rows available after aggregation and suppression.")
} else if (has_crosstalk) {
  trend_shared <- crosstalk::SharedData$new(
    trajectory_trend_summary,
    key = ~paste(trajectory_var, study, clyr, trajectory_class),
    group = "trajectory_trend_group"
  )

  crosstalk::bscols(
    widths = c(2, 2, 2, 2, 4),
    crosstalk::filter_select("traj_dim", "Trajectory Dimension", trend_shared, ~trajectory_label, multiple = TRUE),
    crosstalk::filter_select("traj_study", "Study", trend_shared, ~study, multiple = TRUE),
    crosstalk::filter_select("traj_class", "Class", trend_shared, ~trajectory_class, multiple = TRUE),
    crosstalk::filter_select("traj_clyr", "CLYR", trend_shared, ~clyr, multiple = TRUE),
    datatable_safe(
      trend_shared,
      rownames = FALSE,
      options = list(pageLength = 12, autoWidth = TRUE, scrollX = TRUE),
      colnames = c(
        "Trajectory Variable",
        "Trajectory Dimension",
        "Study",
        "CLYR",
        "Class",
        "N",
        "Depressed N",
        "Depression Prev (%)"
      )
    )
  )
} else {
  datatable_safe(
    trajectory_trend_summary,
    rownames = FALSE,
    options = list(pageLength = 12, autoWidth = TRUE, scrollX = TRUE),
    colnames = c(
      "Trajectory Variable",
      "Trajectory Dimension",
      "Study",
      "CLYR",
      "Class",
      "N",
      "Depressed N",
      "Depression Prev (%)"
    )
  )
}
```

```{r}
#| echo: false
if (nrow(trajectory_pairwise_summary) == 0) {
  cat("No trajectory cross-class rows available. At least two trajectory dimensions with data are required.")
} else if (has_crosstalk) {
  pair_shared <- crosstalk::SharedData$new(
    trajectory_pairwise_summary,
    key = ~paste(pair_id, study, class_x, class_y),
    group = "trajectory_pair_group"
  )

  crosstalk::bscols(
    widths = c(2, 2, 2, 2, 4),
    crosstalk::filter_select("traj_pair", "Trajectory Pair", pair_shared, ~pair_label, multiple = TRUE),
    crosstalk::filter_select("traj_pair_study", "Study", pair_shared, ~study, multiple = TRUE),
    crosstalk::filter_select("traj_pair_x", "Class A", pair_shared, ~class_x, multiple = TRUE),
    crosstalk::filter_select("traj_pair_y", "Class B", pair_shared, ~class_y, multiple = TRUE),
    datatable_safe(
      pair_shared,
      rownames = FALSE,
      options = list(pageLength = 12, autoWidth = TRUE, scrollX = TRUE),
      colnames = c(
        "Pair ID",
        "Trajectory Pair",
        "Trajectory Var A",
        "Trajectory Label A",
        "Class A",
        "Trajectory Var B",
        "Trajectory Label B",
        "Class B",
        "Study",
        "N",
        "Depressed N",
        "Depression Prev (%)"
      )
    )
  )
} else {
  datatable_safe(
    trajectory_pairwise_summary,
    rownames = FALSE,
    options = list(pageLength = 12, autoWidth = TRUE, scrollX = TRUE),
    colnames = c(
      "Pair ID",
      "Trajectory Pair",
      "Trajectory Var A",
      "Trajectory Label A",
      "Class A",
      "Trajectory Var B",
      "Trajectory Label B",
      "Class B",
      "Study",
      "N",
      "Depressed N",
      "Depression Prev (%)"
    )
  )
}
```

### Data

#### Filter Dimension Availability

```{r}
#| echo: false
schema_tbl <- filter_schema
schema_tbl$available <- ifelse(schema_tbl$available, "Yes", "No")
schema_tbl$source_present <- ifelse(schema_tbl$source_present, "Yes", "No")

datatable_safe(
  schema_tbl[, c("domain", "label", "var", "type", "source_present", "available", "distinct_levels", "non_missing_n")],
  rownames = FALSE,
  options = list(pageLength = 15, autoWidth = TRUE, scrollX = TRUE),
  colnames = c(
    "Domain",
    "Label",
    "Variable",
    "Type",
    "Present in Source",
    "Usable in Dashboard",
    "Distinct Levels",
    "Non-missing N"
  )
)
```

#### Demographic Breakdown Table

```{r}
#| echo: false
if (has_crosstalk) {
  demog_shared <- crosstalk::SharedData$new(
    demog_summary,
    key = ~paste(study, sex, age_band),
    group = "demog_group"
  )

  crosstalk::bscols(
    widths = c(2, 2, 2, 6),
    crosstalk::filter_select("demog_study", "Study", demog_shared, ~study, multiple = TRUE),
    crosstalk::filter_select("demog_sex", "Sex", demog_shared, ~sex, multiple = TRUE),
    crosstalk::filter_select("demog_age", "Age Band", demog_shared, ~age_band, multiple = TRUE),
    datatable_safe(
      demog_shared,
      rownames = FALSE,
      options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
      colnames = c("Study", "Sex", "Age Band", "N", "Depressed N", "Depression Prev (%)")
    )
  )
} else {
  datatable_safe(
    demog_summary,
    rownames = FALSE,
    options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
    colnames = c("Study", "Sex", "Age Band", "N", "Depressed N", "Depression Prev (%)")
  )
}
```

#### Visit-Year Table

```{r}
#| echo: false
if (has_crosstalk) {
  year_shared <- crosstalk::SharedData$new(
    year_summary,
    key = ~paste(study, clyr),
    group = "year_group"
  )

  crosstalk::bscols(
    widths = c(3, 9),
    crosstalk::filter_select("year_study", "Study", year_shared, ~study, multiple = TRUE),
    datatable_safe(
      year_shared,
      rownames = FALSE,
      options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
      colnames = c("Study", "CLYR", "N", "Depressed N", "Depression Prev (%)")
    )
  )
} else {
  datatable_safe(
    year_summary,
    rownames = FALSE,
    options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
    colnames = c("Study", "CLYR", "N", "Depressed N", "Depression Prev (%)")
  )
}
```

#### Build Metadata

```{r}
#| echo: false
meta_tbl <- data.frame(
  Field = c(
    "Build timestamp",
    "Input rows (private source, not stored)",
    "Analysis rows after DPBN filter",
    "Participant snapshot rows",
    "Small-cell suppression threshold",
    "Configured filter dimensions",
    "Available filter dimensions",
    "Filter cube rows",
    "Trajectory dimensions available",
    "Trajectory trend rows",
    "Trajectory cross-class rows"
  ),
  Value = c(
    dashboard_build_info$rendered_at,
    dashboard_build_info$n_input_rows,
    dashboard_build_info$n_analysis_rows,
    dashboard_build_info$n_participant_snapshot,
    dashboard_build_info$min_cell_n,
    dashboard_build_info$n_filter_dimensions_configured,
    dashboard_build_info$n_filter_dimensions_available,
    dashboard_build_info$n_filter_cube_rows,
    dashboard_build_info$n_trajectory_dimensions_available,
    dashboard_build_info$n_trajectory_trend_rows,
    dashboard_build_info$n_trajectory_pairwise_rows
  ),
  check.names = FALSE
)

knitr::kable(meta_tbl, align = c("l", "l"), row.names = FALSE)
```
