---
title: "**Dashboard**"
format:
  dashboard:
    theme: sandstone
    css: ../styles.css
toc: false
---

```{r}
#| include: false
#| warning: false
#| message: false
read_obj <- function(filename) {
  candidates <- c(
    file.path("_tools", "_objects", filename),
    file.path("..", "_tools", "_objects", filename)
  )
  idx <- which(file.exists(candidates))[1]
  if (is.na(idx)) {
    stop("Missing object file: ", filename)
  }
  readRDS(candidates[idx])
}

kpi <- read_obj("dashboard_kpi.rds")

fmt_int <- function(x) format(round(x), big.mark = ",", trim = TRUE)
fmt_pct <- function(x) ifelse(is.na(x), "NA", sprintf("%.1f%%", x))

mh_start_label <- if (!is.null(kpi$mh_start_month_year) && nzchar(kpi$mh_start_month_year)) {
  kpi$mh_start_month_year
} else {
  "Not available"
}
mh_end_label <- if (!is.null(kpi$mh_end_month_year) && nzchar(kpi$mh_end_month_year)) {
  kpi$mh_end_month_year
} else {
  "Not available"
}
mh_range_label <- if (!is.null(kpi$mh_range_month_year) && nzchar(kpi$mh_range_month_year)) {
  kpi$mh_range_month_year
} else {
  "Not available"
}
sdoh_start_label <- if (!is.null(kpi$sdoh_start_year) && nzchar(kpi$sdoh_start_year)) {
  kpi$sdoh_start_year
} else {
  "Not available"
}
sdoh_end_label <- if (!is.null(kpi$sdoh_end_year) && nzchar(kpi$sdoh_end_year)) {
  kpi$sdoh_end_year
} else {
  "Not available"
}
sdoh_range_label <- if (!is.null(kpi$sdoh_range_year) && nzchar(kpi$sdoh_range_year)) {
  kpi$sdoh_range_year
} else {
  "Not available"
}
sdoh_start_my_label <- if (!is.null(kpi$sdoh_start_month_year) && nzchar(kpi$sdoh_start_month_year)) {
  kpi$sdoh_start_month_year
} else if (sdoh_start_label != "Not available") {
  paste("Jan", sdoh_start_label)
} else {
  "Not available"
}
sdoh_end_my_label <- if (!is.null(kpi$sdoh_end_month_year) && nzchar(kpi$sdoh_end_month_year)) {
  kpi$sdoh_end_month_year
} else if (sdoh_end_label != "Not available") {
  paste("Dec", sdoh_end_label)
} else {
  "Not available"
}
sdoh_range_my_label <- if (!is.null(kpi$sdoh_range_month_year) && nzchar(kpi$sdoh_range_month_year)) {
  kpi$sdoh_range_month_year
} else if (sdoh_start_my_label != "Not available" && sdoh_end_my_label != "Not available") {
  paste0(sdoh_start_my_label, " to ", sdoh_end_my_label)
} else {
  "Not available"
}
```

# {.sidebar}

::: {style="text-align: center;"}
**Dataset Coverage**

Mental health (single-timepoint)  
*`r mh_range_label`*

SDoH trajectories (longitudinal)  
*`r sdoh_range_my_label`*

<p style="margin-top: 1rem; font-size: 0.9em;"><em>Disclaimer: This dashboard uses aggregated data only.</em></p>
:::

# All {orientation="rows"}

## Row {height="8%"}

### Column {width="15%"}

```{r}
#| echo: false
#| content: valuebox
#| title: "Cohorts"
list(
  icon = "diagram-3",
  color = "info",
  value = fmt_int(kpi$studies_total)
)
```

### Column {width="25.5%"}

```{r}
#| echo: false
#| content: valuebox
#| title: "Participants/Records"
list(
  icon = "people-fill",
  color = "primary",
  value = paste(fmt_int(kpi$participants_total), "/", fmt_int(kpi$records_total))
)
```

### Column {width="25.5%"}

```{r}
#| echo: false
#| content: valuebox
#| title: "Depression (%)"
list(
  icon = "activity",
  color = "warning",
  value = fmt_pct(kpi$depression_prev_pct)
)
```

### Column {width="25.5%"}

```{r}
#| echo: false
#| content: valuebox
#| title: "Subgroups"
list(
  icon = "funnel",
  color = "primary",
  value = fmt_int(kpi$filter_dimensions_available)
)
```

### Column {width="27.5%"}

```{r}
#| echo: false
#| content: valuebox
#| title: "Trajectory dimensions"
list(
  icon = "bezier2",
  color = "secondary",
  value = fmt_int(kpi$trajectory_dimensions_available)
)
```

## Row {height="46%"}

### Column {width="40%"}

```{r}
#| title: Plot
```

### Column {width="60%"}

```{r}
#| title: Plot
```

## Row {height="46%"}

```{r}
#| title: Plot
```

```{r}
#| title: Plot
```

# SDoH

## Row {height="40%"}

### Column {width="40%"}

```{r}
#| title: Plot
```

### Column {width="60%"}

```{r}
#| title: Plot
```

## Row {height="40%"}

```{r}
#| title: Plot
```

```{r}
#| title: Plot
```


# Trajectories

<!-- Intentionally empty for now. -->

# Data

<!-- Intentionally empty for now. -->
