---
title: "Longitudinal Linkage"
toc: true
---

## Purpose

This page documents the longitudinal linkage stage that combines mental health outcome harmonization and SDoH trajectory information into the final integrated secondary analysis dataset.

## Goal Alignment

The linkage workflow directly supports the project objectives:

1. Build and maintain a linked analysis dataset.
2. Quantify direct and mediation relationships for depression outcomes.
3. Characterize social determinant trajectories across childhood windows.

## Stage Flow (Current Build Counts)

```{r}
#| include: false

find_project_root <- function(start_dir = normalizePath(getwd(), winslash = "/", mustWork = TRUE)) {
  dir_now <- start_dir
  repeat {
    if (file.exists(file.path(dir_now, "_quarto.yml"))) {
      return(dir_now)
    }
    parent <- dirname(dir_now)
    if (identical(parent, dir_now)) {
      return(start_dir)
    }
    dir_now <- parent
  }
}

safe_load_object <- function(rdata_path, object_name) {
  if (!file.exists(rdata_path)) {
    return(NULL)
  }
  e <- new.env(parent = emptyenv())
  load(rdata_path, envir = e)
  if (!exists(object_name, envir = e, inherits = FALSE)) {
    return(NULL)
  }
  get(object_name, envir = e, inherits = FALSE)
}

count_distinct_ids <- function(df, id_var = "USUBJID", complete_vars = NULL, require_non_missing = NULL) {
  if (is.null(df) || !(id_var %in% names(df))) {
    return(NA_integer_)
  }

  keep <- !is.na(df[[id_var]])

  if (!is.null(complete_vars)) {
    vars <- intersect(complete_vars, names(df))
    if (length(vars) > 0) {
      keep <- keep & stats::complete.cases(df[, vars, drop = FALSE])
    }
  }

  if (!is.null(require_non_missing) && (require_non_missing %in% names(df))) {
    keep <- keep & !is.na(df[[require_non_missing]])
  }

  length(unique(df[[id_var]][keep]))
}

fmt_n <- function(x) {
  if (is.na(x)) {
    return("?")
  }
  format(x, big.mark = ",", trim = TRUE, scientific = FALSE)
}

project_root <- find_project_root()
private_subpath <- file.path("_private_use", "data_management", "data", "co-luminate", "adam")
private_candidates <- c(
  file.path(dirname(dirname(project_root)), private_subpath),
  file.path(project_root, "..", "..", private_subpath),
  file.path(project_root, "..", private_subpath)
)
private_candidates <- unique(normalizePath(private_candidates, winslash = "/", mustWork = FALSE))
private_adam_dir <- private_candidates[
  file.exists(file.path(private_candidates, "dt_multistudies.RData"))
][1]

if (is.na(private_adam_dir) || is.null(private_adam_dir)) {
  private_adam_dir <- ""
}

psychometric_df <- safe_load_object(file.path(private_adam_dir, "dt_psychometric.RData"), "dt_psychometric")
ssq_only_df <- safe_load_object(file.path(private_adam_dir, "dt_dreams_multi_ssq.RData"), "dt_dreams_multi_ssq")
childhood_df <- safe_load_object(file.path(private_adam_dir, "dt_childhood_exposure.RData"), "dt_childhood_exposure")
integrated_df <- safe_load_object(file.path(private_adam_dir, "dt_multistudies.RData"), "dt_multistudies")

calibration_vars <- c("PHQSCR", "SSQ01", "SSQ02", "SSQ03", "SSQ08", "SSQ09", "SSQ10", "SSQ11", "SSQ12", "SSQ13", "SSQ14")

n_a <- count_distinct_ids(psychometric_df)
n_b <- count_distinct_ids(psychometric_df, complete_vars = calibration_vars)
n_c <- count_distinct_ids(ssq_only_df)
n_d <- count_distinct_ids(ssq_only_df, require_non_missing = "TDEPBIN")
n_e <- count_distinct_ids(childhood_df)

n_ids_05 <- character(0)
n_ids_612 <- character(0)

n_f_05 <- {
  if (is.null(childhood_df) || !("USUBJID" %in% names(childhood_df)) || !("EXPAGE" %in% names(childhood_df))) {
    NA_integer_
  } else {
    keep <- !is.na(childhood_df$USUBJID) &
      !is.na(childhood_df$EXPAGE) &
      childhood_df$EXPAGE >= 0 &
      childhood_df$EXPAGE <= 5
    n_ids_05 <<- unique(as.character(childhood_df$USUBJID[keep]))
    length(n_ids_05)
  }
}

n_f_612 <- {
  if (is.null(childhood_df) || !("USUBJID" %in% names(childhood_df)) || !("EXPAGE" %in% names(childhood_df))) {
    NA_integer_
  } else {
    keep <- !is.na(childhood_df$USUBJID) &
      !is.na(childhood_df$EXPAGE) &
      childhood_df$EXPAGE >= 6 &
      childhood_df$EXPAGE <= 12
    n_ids_612 <<- unique(as.character(childhood_df$USUBJID[keep]))
    length(n_ids_612)
  }
}

n_f_both <- if (is.na(n_f_05) || is.na(n_f_612)) NA_integer_ else length(intersect(n_ids_05, n_ids_612))
n_f_only_05 <- if (is.na(n_f_05) || is.na(n_f_612)) NA_integer_ else length(setdiff(n_ids_05, n_ids_612))
n_f_only_612 <- if (is.na(n_f_05) || is.na(n_f_612)) NA_integer_ else length(setdiff(n_ids_612, n_ids_05))

n_g <- count_distinct_ids(integrated_df, require_non_missing = "DPBN")
sdoh_component_ids <- if (is.na(n_f_05) || is.na(n_f_612)) character(0) else union(n_ids_05, n_ids_612)
dpbn_endpoint_ids <- {
  if (is.null(integrated_df) || !("USUBJID" %in% names(integrated_df)) || !("DPBN" %in% names(integrated_df))) {
    character(0)
  } else {
    unique(as.character(integrated_df$USUBJID[!is.na(integrated_df$USUBJID) & !is.na(integrated_df$DPBN)]))
  }
}

n_h <- if (length(sdoh_component_ids) == 0) NA_integer_ else length(sdoh_component_ids)
n_i <- if (length(sdoh_component_ids) == 0 || length(dpbn_endpoint_ids) == 0) {
  NA_integer_
} else {
  length(intersect(sdoh_component_ids, dpbn_endpoint_ids))
}
```

```{r}
#| echo: false
#| results: asis

diagram_lines <- c(
  "```{mermaid}",
  "%%{init: {\"flowchart\": {\"curve\": \"linear\", \"htmlLabels\": true}} }%%",
  "flowchart LR",
  "  O1[\"Objective 1<br/>Build linked<br/>analysis dataset\"]",
  "  O2[\"Objective 2<br/>Quantify direct<br/>and mediation links\"]",
  "  O3[\"Objective 3<br/>Characterize SDoH<br/>trajectories\"]",
  "",
  "  S1[\"Linkage Stage 1<br/>Psychometric<br/>calibration\"]",
  "  S2[\"Linkage Stage 2<br/>SDoH trajectory<br/>derivation\"]",
  "  S3[\"Linkage Stage 3<br/>Final linkage<br/>and integration\"]",
  "",
  "  O1 --> S1",
  "  O1 --> S2",
  "  O1 --> S3",
  "  O2 --> S3",
  "  O3 --> S2",
  "",
  sprintf("  S1 --> A[\"Psychometric bridge<br/>source data<br/>(n=%s)\"]", fmt_n(n_a)),
  sprintf("  A --> B[\"Fit IRT calibration<br/>model (SSQ to PHQ)<br/>reference scale<br/>(n=%s)\"]", fmt_n(n_b)),
  sprintf("  S1 --> C[\"SSQ-only cohort<br/>response data<br/>(n=%s)\"]", fmt_n(n_c)),
  sprintf("  C --> D[\"Apply calibration<br/>for harmonized<br/>depression endpoint<br/>(n=%s)\"]", fmt_n(n_d)),
  "  B --> D",
  "",
  sprintf("  S2 --> E[\"Childhood exposure<br/>longitudinal data<br/>(n=%s)\"]", fmt_n(n_e)),
  sprintf("  E --> F[\"Derive SDoH<br/>trajectories<br/>0-5 years (n=%s)<br/>6-12 years (n=%s)<br/>Both (n=%s)<br/>Only 0-5 (n=%s)<br/>Only 6-12 (n=%s)<br/><br/>\"]", fmt_n(n_f_05), fmt_n(n_f_612), fmt_n(n_f_both), fmt_n(n_f_only_05), fmt_n(n_f_only_612)),
  "",
  sprintf("  S3 --> G[\"Assemble mental<br/>health component<br/>(n=%s)\"]", fmt_n(n_g)),
  sprintf("  S3 --> H[\"Assemble SDoH<br/>trajectory component<br/>(n=%s)\"]", fmt_n(n_h)),
  "  D --> G",
  "  F --> H",
  "",
  sprintf("  G --> I[\"Integrated secondary<br/>analysis dataset<br/>(n=%s)\"]", fmt_n(n_i)),
  "  H --> I",
  "",
  "  classDef objective fill:#eef2ff,stroke:#4338ca,color:#1e1b4b,stroke-width:1.5px;",
  "  classDef stage1 fill:#dbeafe,stroke:#1d4ed8,color:#0f172a,stroke-width:1.5px;",
  "  classDef stage2 fill:#fff7ed,stroke:#c2410c,color:#431407,stroke-width:1.5px;",
  "  classDef stage3 fill:#ecfdf5,stroke:#059669,color:#064e3b,stroke-width:1.5px;",
  "  classDef psych fill:#dbeafe,stroke:#1d4ed8,color:#0f172a;",
  "  classDef sdoh fill:#ffedd5,stroke:#c2410c,color:#431407;",
  "  classDef integ fill:#dcfce7,stroke:#059669,color:#064e3b,stroke-width:1.5px;",
  "  class O1,O2,O3 objective;",
  "  class S1 stage1;",
  "  class S2 stage2;",
  "  class S3 stage3;",
  "  class A,B,C,D,G psych;",
  "  class E,F,H sdoh;",
  "  class I integ;",
  "```"
)

cat(paste(diagram_lines, collapse = "\n"))
```

## Interpretation

1. Stage 1 calibrates depression information for SSQ-only records.
2. Stage 2 derives SDoH trajectory coverage across 0-5 and 6-12 windows.
3. Stage 3 integrates both streams and keeps participants with both:
`Depression endpoint` and `SDoH trajectory coverage`.

## Completion Criteria

This stage is complete when:

1. Depression endpoint values are available for integrated mental health records.
2. SDoH trajectory coverage is available across childhood windows.
3. Final integrated data retains only participants meeting both criteria.

## Notes

- Counts are distinct participants (`USUBJID`) and are calculated at render time.
- Violet nodes: proposal objectives.
- Blue nodes: psychometric stage.
- Orange nodes: SDoH trajectory stage.
- Green nodes: final integration stage.
