---
title: "Metadata"
toc: true
---

## Metadata Source

This page is extracted and adapted from the dataset preparation source index.

Current scope is limited to two data dictionaries:

- Childhood social exposure panel
- Integrated multi-study mental health panel

All derived R objects for this page are stored in `PRJ-CO-LUMINATE/_tools/_objects`.

```{r}
#| echo: false
read_obj <- function(filename) {
  candidates <- c(
    file.path("_tools", "_objects", filename),
    file.path("..", "_tools", "_objects", filename)
  )
  idx <- which(file.exists(candidates))[1]
  if (is.na(idx)) {
    stop("Missing object file: ", filename)
  }
  readRDS(candidates[idx])
}

render_role_dictionary <- function(role_df) {
  role_df <- role_df[, c("analysis_role", "variable", "label", "type", "value_label"), drop = FALSE]
  names(role_df) <- c("Role", "Variable", "Label", "Type", "Value Labels")

  if (requireNamespace("gt", quietly = TRUE)) {
    return(
      gt::gt(role_df, groupname_col = "Role") |>
        gt::cols_hide(columns = "Role") |>
        gt::tab_style(
          style = gt::cell_text(weight = "bold"),
          locations = gt::cells_row_groups()
        )
    )
  }

  role_df$Role <- ifelse(
    duplicated(role_df$Role),
    "",
    paste0("**", role_df$Role, "**")
  )

  knitr::kable(
    role_df,
    align = c("l", "l", "l", "l", "l"),
    escape = FALSE,
    row.names = FALSE
  )
}

append_missing_to_role_dictionary <- function(role_df, missing_df) {
  if (is.null(missing_df) || nrow(missing_df) == 0) {
    return(role_df)
  }

  missing_df <- missing_df[!missing_df$variable %in% role_df$variable, , drop = FALSE]
  if (nrow(missing_df) == 0) {
    return(role_df)
  }

  add_rows <- data.frame(
    analysis_role = missing_df$analysis_role,
    variable = missing_df$variable,
    label = "Configured variable",
    type = "",
    value_label = "Missing in current object",
    stringsAsFactors = FALSE
  )

  rbind(role_df, add_rows)
}
```

## Dataset Summary

```{r}
#| echo: false
metadata_summary <- read_obj("metadata_summary.rds")
knitr::kable(
  metadata_summary,
  col.names = c("Field", "Value"),
  align = c("l", "l"),
  row.names = FALSE
)
```

## Study Design Extract

The CO-LUMINATE linked dataset connects individual-level mental health outcomes from cohort studies to household-level longitudinal HDSS records. The objective is to estimate longitudinal direct and mediation relationships between social determinants and depressive symptoms among youth aged 13-24.

The analytical design combines:

1. HDSS longitudinal household exposure histories.
2. Cohort-based depression outcomes measured with SSQ-14 and PHQ-9.
3. Linkage across five cohorts nested within AHRI HDSS.

## Cohorts Included

- DREAMS
- Multilevel
- Isisekelo Sempilo
- Treatment as Prevention (TasP)
- Thetha Nami

## Object Snapshot

```{r}
#| echo: false
object_snapshot <- read_obj("object_snapshot.rds")

knitr::kable(object_snapshot, align = c("l", "r", "r"), row.names = FALSE)
```

## CONSORT Flow

```{r}
#| echo: false
consort_counts <- read_obj("consort_counts.rds")
consort_exclusion_studies <- read_obj("consort_exclusion_studies.rds")

metric_value <- function(metric) {
  idx <- match(metric, consort_counts$metric)
  if (is.na(idx)) return(NA_integer_)
  consort_counts$value[idx]
}

fmt_n <- function(x) {
  if (is.na(x)) return("NA")
  format(as.integer(x), big.mark = ",", trim = TRUE, scientific = FALSE)
}

n_multi_loaded <- metric_value("multi_participants_loaded")
n_multi_with_dpbn <- metric_value("multi_participants_with_dpbn")
n_child_loaded <- metric_value("child_participants_loaded")
n_overlap <- metric_value("overlap_participants")
n_final <- metric_value("multi_participants_final")

n_excl_missing_dpbn <- if (is.na(n_multi_loaded) || is.na(n_multi_with_dpbn)) NA_integer_ else n_multi_loaded - n_multi_with_dpbn
n_excl_multi_no_sdoh <- if (is.na(n_multi_with_dpbn) || is.na(n_overlap)) NA_integer_ else n_multi_with_dpbn - n_overlap
n_excl_child_no_multi <- if (is.na(n_child_loaded) || is.na(n_overlap)) NA_integer_ else n_child_loaded - n_overlap
n_mh_remaining_after_endpoint <- n_multi_with_dpbn
n_mh_remaining_after_match <- n_overlap
n_sdoh_remaining_after_match <- if (is.na(n_child_loaded) || is.na(n_excl_child_no_multi)) NA_integer_ else n_child_loaded - n_excl_child_no_multi

user_lib <- Sys.getenv("R_LIBS_USER", unset = "")
if (nzchar(user_lib) && dir.exists(user_lib)) {
  .libPaths(unique(c(user_lib, .libPaths())))
}

study_value <- function(exclusion_key) {
  idx <- match(exclusion_key, consort_exclusion_studies$exclusion_key)
  if (is.na(idx)) return("Study source unavailable")
  if ("study_breakdown" %in% names(consort_exclusion_studies)) {
    return(consort_exclusion_studies$study_breakdown[idx])
  }
  consort_exclusion_studies$studies[idx]
}

fmt_study_lines <- function(x) {
  if (is.na(x) || !nzchar(x)) {
    return("Study source unavailable")
  }
  split_pat <- if (grepl(";", x, fixed = TRUE)) ";" else ","
  parts <- trimws(strsplit(x, split_pat, fixed = TRUE)[[1]])
  parts <- parts[nzchar(parts)]
  if (length(parts) == 0) return("Study source unavailable")
  paste(parts, collapse = "<br/>")
}

if (requireNamespace("DiagrammeR", quietly = TRUE)) {
  mermaid_code <- paste(
    "%%{init: {\"themeVariables\": {\"fontSize\": \"20px\"}, \"flowchart\": {\"nodeSpacing\": 80, \"rankSpacing\": 110}}}%%",
    "graph TD",
    sprintf("A[\"Mental health<br/>available<br/>(n=%s)\"]", fmt_n(n_multi_loaded)),
    sprintf(
      "X1[\"Excluded<br/>Missing depression<br/>endpoint (n=%s)<br/>Studies:<br/>%s\"]",
      fmt_n(n_excl_missing_dpbn),
      fmt_study_lines(study_value("missing_depression_endpoint"))
    ),
    sprintf("B[\"Mental health<br/>remaining after<br/>endpoint filter<br/>(n=%s)\"]", fmt_n(n_mh_remaining_after_endpoint)),
    sprintf(
      "X2[\"Excluded<br/>No SDoH match<br/>(n=%s)<br/>Studies:<br/>%s\"]",
      fmt_n(n_excl_multi_no_sdoh),
      fmt_study_lines(study_value("no_sdoh_match"))
    ),
    sprintf("M[\"Mental health<br/>remaining after<br/>match filter<br/>(n=%s)\"]", fmt_n(n_mh_remaining_after_match)),
    sprintf("C[\"SDoH<br/>available<br/>(n=%s)\"]", fmt_n(n_child_loaded)),
    sprintf(
      "X3[\"Excluded<br/>No eligible mental<br/>health match<br/>(n=%s)<br/>Studies:<br/>%s\"]",
      fmt_n(n_excl_child_no_multi),
      fmt_study_lines(study_value("no_eligible_mental_health_match"))
    ),
    sprintf("S[\"SDoH remaining<br/>after match filter<br/>(n=%s)\"]", fmt_n(n_sdoh_remaining_after_match)),
    sprintf("D[\"Final included<br/>Depression endpoint<br/>and SDoH coverage<br/>(n=%s)\"]", fmt_n(n_final)),
    "A --> B",
    "B --> M",
    "M --> D",
    "C --> S",
    "S --> D",
    "A --> X1",
    "B --> X2",
    "C --> X3",
    "classDef included fill:#ecfdf5,stroke:#059669,color:#064e3b,stroke-width:1.5px;",
    "classDef excluded fill:#fff1f2,stroke:#e11d48,color:#4c0519,stroke-width:1.5px;",
    "class A,B,M,C,S,D included;",
    "class X1,X2,X3 excluded;",
    sep = "\n"
  )
  DiagrammeR::mermaid(mermaid_code, width = "100%", height = "900px")
} else {
  cat("DiagrammeR package is not available in this environment.")
}
```

Participant-focused CONSORT logic:
participants are retained only if they have both a depression endpoint and SDoH trajectory coverage.

The row/column snapshot above is based on the current metadata object build inputs. Dictionary definitions below are extracted from the dataset preparation master dictionary specification.

## Data Dictionary: Childhood Social Exposure Panel (Variables of Interest by Role)

```{r}
#| echo: false
dt_chilhood_exposure_role_dictionary <- read_obj("dt_chilhood_exposure_role_dictionary.rds")
dt_chilhood_exposure_role_missing <- read_obj("dt_chilhood_exposure_role_missing.rds")
dt_chilhood_exposure_with_missing <- append_missing_to_role_dictionary(
  dt_chilhood_exposure_role_dictionary,
  dt_chilhood_exposure_role_missing
)
render_role_dictionary(dt_chilhood_exposure_with_missing)
```

## Data Dictionary: Integrated Multi-study Mental Health Panel (Variables of Interest by Role)

```{r}
#| echo: false
dt_multistudies_role_dictionary <- read_obj("dt_multistudies_role_dictionary.rds")
render_role_dictionary(dt_multistudies_role_dictionary)
```

## Data Dictionary: Integrated Multi-study Mental Health Panel (Derived Variables)

```{r}
#| echo: false
dt_multistudies_derived_dictionary <- read_obj("dt_multistudies_derived_dictionary.rds")
dt_multistudies_derived_dictionary <- dt_multistudies_derived_dictionary[
  !dt_multistudies_derived_dictionary$variable %in% c("CLYR", "GBCDV"),
  ,
  drop = FALSE
]
if ("derivation_note" %in% names(dt_multistudies_derived_dictionary)) {
  dt_multistudies_derived_dictionary$derivation_note <- gsub(
    "\\b[A-Z]{3}[0-9]{2}\\b",
    "dataset preparation",
    dt_multistudies_derived_dictionary$derivation_note
  )
}
if ("derivation_status" %in% names(dt_multistudies_derived_dictionary)) {
  dt_multistudies_derived_dictionary$derivation_status <- gsub(
    "\\b[A-Z]{3}[0-9]{2}\\b",
    "dataset preparation",
    dt_multistudies_derived_dictionary$derivation_status
  )
}
if ("DPBN" %in% dt_multistudies_derived_dictionary$variable) {
  dpbn_idx <- which(dt_multistudies_derived_dictionary$variable == "DPBN")[1]
  dt_multistudies_derived_dictionary$source_variables[dpbn_idx] <- "PHQBIN, irt_joint_models.rds"
  dt_multistudies_derived_dictionary$derivation_note[dpbn_idx] <- "PHQBIN >= 10"
  dt_multistudies_derived_dictionary$derivation_status[dpbn_idx] <- "Implemented in dataset preparation merge"
}
if (nrow(dt_multistudies_derived_dictionary) == 0) {
  cat("No derived variables were identified in the current metadata objects.")
} else {
  knitr::kable(
    dt_multistudies_derived_dictionary,
    col.names = c(
      "Variable",
      "Label",
      "Type",
      "Value Labels",
      "Source Variables",
      "Derivation",
      "Status"
    ),
    align = c("l", "l", "l", "l", "l", "l", "l"),
    row.names = FALSE
  )
}
```

## Dataset Name Reference

```{r}
#| echo: false
dataset_name_reference <- data.frame(
  Dataset_Name = c(
    "dt_childhood_exposure",
    "dt_multistudies",
    "dt_psychometric",
    "dt_dreams_multi_ssq"
  ),
  Readable_Name = c(
    "Childhood Social Exposure Panel",
    "Integrated Multi-study Mental Health Panel",
    "Psychometric Bridge Source Data",
    "SSQ-only Cohort Response Data"
  ),
  stringsAsFactors = FALSE
)

knitr::kable(
  dataset_name_reference,
  col.names = c("Dataset Name", "Readable Name"),
  align = c("l", "l"),
  row.names = FALSE
)
```
