---
title: Establishing Comparability between the Shona Symptom Questionnaire (SSQ-14) and the Patient Health Questionnaire (PHQ-9) as Depression Outcome Measures among Adolescents and Young Adults in South Africa
subtitle: "**Statistical Consideration and Analysis Report (SCAR)**"

authors:
  - name: CO-LUMINATE Research Team
    affiliation: Africa Health Research Institute (AHRI)
    city: Durban & Somokhele
    country: South Africa
    
bibliography: references.bib
---

Word Count:

Impact Statement: Harmonising mental health data from African adolescents enables the robust longitudinal analysis required to inform policy, yet fragmentation between indigenous and global measures impedes this progress. This research provides a methodological roadmap for linking the Shona Symptom Questionnaire and the Patient Health Questionnaire, unlocking the potential of existing datasets to address the youth mental health crisis in Southern Africa.

Abstract
Fragmented measurement strategies in mental health research significantly reduce the utility of pooled datasets, particularly in the Global South where data scarcity is acute. This paper addresses the methodological challenge of establishing comparability between an emic, indigenous measure of distress, the Shona Symptom Questionnaire (SSQ-14), and a global etic standard, the Patient Health Questionnaire (PHQ-9). We review the imperatives for harmonisation, emphasising the statistical power required to study low-prevalence outcomes and the causal inference benefits of longitudinal designs. We critique the current measurement landscape in South Africa, which is characterised by "measurement silos" that prevent the integration of valuable cohort data. Special attention is given to the pitfalls of traditional Receiver Operating Characteristic (ROC) analysis and binary cut-off scores, arguing instead for methods that preserve continuous variation. Utilising the Maelstrom Research Guidelines, we delineate a framework for retrospective harmonisation that addresses structural divergences such as differing recall periods and response formats. We argue that Item Response Theory (IRT) provides the most robust solution for linking the binary items of the SSQ-14 with the ordinal items of the PHQ-9 on a common latent trait of general distress. This approach, supported by recent validation studies in Zanzibar, offers a pathway to support data sovereignty and advance equitable mental health science in Africa.

Keywords: Data Harmonisation, Mental Health, SSQ-14, PHQ-9, South Africa, Psychometrics

## Strategic Implications for African Mental Health Science
### Advancing Causal Inference in Longitudinal Models
Once harmonised, SSQ-14 and PHQ-9 scores can be treated as a single continuous variable in longitudinal models. This allows for the use of fixed-effects regression and growth-curve modeling to track distress trajectories from adolescence to young adulthood, even if the cohort switched from the SSQ-14 to the PHQ-9 over time. The harmonisation bridge allows the slope of the curve to be estimated without a discontinuity artifact at the point of instrument switch.

### Supporting Data Sovereignty
By establishing these harmonisation algorithms locally using African validation studies, we empower African researchers to lead the analysis of their own data. Tools like DataSHIELD can be deployed with these harmonisation scripts embedded, allowing for federated analysis of mental health across the continent without the data ever leaving African servers [@gaye2014]. This fulfills the vision of initiatives like the H3Africa Consortium for equitable, high-quality, and locally governed science [@mulder2021].

## Conclusion

The harmonisation of the SSQ-14 and PHQ-9 is a challenging but technically feasible endeavor that is essential for the future of mental health research in South Africa. While the instruments emerge from different paradigms—one emic and binary, the other etic and polytomous—they share a substantial core of somatic and affective content that serves as a bridge for comparability. The recommended pathway involves a rigorous adherence to the Maelstrom Guidelines: precise construct definition, careful content mapping to identify anchor items, and the application of Item Response Theory to create a sample-independent cross-walk. By bridging these measures, we do more than fix a statistical problem; we unite two traditions of understanding mental illness. This harmonization will unlock the power of existing longitudinal datasets, enabling the robust causal modeling required to identify the determinants of mental health for Africa's youth and to design interventions that are both evidence-based and culturally resonant.

```{r setup, include=FALSE}
#| echo: false
## This chunk handles all package installations and loading for the factor analysis.
require(here)
### Load Helper Functions------------
source(here("../../pipelines_and_functions/data_processing_and_description_funcs.R"))

renv::snapshot()
packages <- c("here", "conflicted", 
              "tidyverse", "labelled", "reshape2",
              "gtsummary", "gt", "gtreg", 
              "scales", "ggsci", "patchwork",
              
              "BiocManager", "htmltools",
              "ComplexUpset","ggstatsplot",
              "psych", "lavaan", "semTools", "pROC",
              "parallel","doParallel","foreach"
              )

## Loop through the packages
func_install_and_load(packages)
renv::snapshot()
conflict_prefer_all("dplyr",quiet=TRUE)
rm(packages)
```

```{r figure-table-setup}
#| echo: false
## This chunk sets global options for figures and tables
theme_gtsummary_compact()
theme_gtsummary_journal(journal = "jama")
options(scipen=999)
```

```{r}
# Create directory for statistical analysis
# Define the root directory name
root_dir <- here("statistical_analysis")

# Define the full list of directories to create
# Use file.path for OS-independent path construction
dirs_to_create <- c(
  file.path(root_dir, "data", "metadata"),
  file.path(root_dir, "data", "adam"),
  file.path(root_dir, "programs/analysis"),
  file.path(root_dir, "programs/utils"),
  file.path(root_dir, "models"),
  file.path(root_dir, "output", "table"),
  file.path(root_dir, "output", "figures"),
  file.path(root_dir, "output", "listing"),
  file.path(root_dir, "output", "objects")
)

# Create all directories recursively
# The 'showWarnings = FALSE' argument prevents warnings if the directories already exist
invisible(sapply(dirs_to_create, dir.create, recursive = TRUE, showWarnings = FALSE))

```

```{r load-data}
#| echo: false
## Load datasets
load(file=here("../../data_management/data/co-luminate/adam/dt_psychometric.RData"))
load(file=here("../../data_management/data/co-luminate/adam/dt_disposition.RData"))
load(file=here("../../data_management/data/co-luminate/adam/dt_concordance.RData"))
load(file=here("../../data_management/data/co-luminate/adam/dt_to_score.RData"))
```

```{r, load-psychometric-efa-pipeline}
# Source the Helper functions for Psychometric Bootstrap EFA Pipeline
source(here("statistical_analysis/programs/utils/Psych Run Pipeline.R"))
```

## Preliminary Analysis

This preliminary analysis evaluates the structural validity and stability of the Shona Symptom Questionnaire (SSQ-14) and two theoretically derived subsets: the SSQ-10 and the SSQ-08. This step serves as a critical foundation for subsequent work, specifically the derivation of an optimal cut-off score for the SSQ-10, referred to as the Theoretically Derived Depressive Symptoms scale, benchmarked against the PHQ-9.

Using a bootstrap resampling framework with 1,000 iterations, we sought to identify a stable factor structure that remains consistent across differing extraction criteria. Rather than viewing the traditional Kaiser Criterion and Parallel Analysis (PA) as competitors, this analysis interprets their results as complementary bounds. The Kaiser Criterion typically favors parsimony, potentially under-extracting, while PA is sensitive to distinct, latent dimensions, potentially extracting minor factors.

```{r, run-psychometric-efa}
# -----------------------------------------------------------------------------
# 7. Run Psychometric Bootstrap EFA Pipeline with Parallelization
# -----------------------------------------------------------------------------

data_name <- "dt_psychometric"
dt_main   <- get(data_name)

n_boot_iterations <- 1000   # Consider testing with smaller first
n_pa_iterations   <- 100
cutoff_value      <- 0.4

item_sets_list <- list(
  "SSQ-14 (Original)"                         = paste0("SSQ", sprintf("%02d", 1:14)),
  "SSQ-10 (Theoretical Depressive Symptoms)" = paste0("SSQ", sprintf("%02d", c(1:3, 8:14))),
  "SSQ-08 (Dixon Chibanda et.al)"           = paste0("SSQ", sprintf("%02d", c(1, 8:14))),
  "PHQ-09 (Original)"                        = paste0("PHQ9", sprintf("%02d", 1:9))
)

main_results <- run_psychometric_pipeline(
  df_data       = dt_main,
  dataset_label = "Isisikelo Sempilo",
  item_sets_list = item_sets_list,
  n_boot        = n_boot_iterations,
  n_pa          = n_pa_iterations,
  loading_cutoff = cutoff_value,
  use_parallel   = TRUE,
  n_cores        = min(10,parallel::detectCores() - 4)
)

output_dir <- here("statistical_analysis/output/objects")
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

saveRDS(main_results,file=file.path(output_dir,
                                    paste0("main_results_boot_",
                                           n_boot_iterations,
                                           "_Isisikelo_Sempilo.rds"))
        )
main_results<-readRDS(file=file.path(output_dir,
                                    paste0("main_results_boot_",
                                           n_boot_iterations,
                                           "_Isisikelo_Sempilo.rds"))
        )
```

### Factor Retention Analysis: Determining the Stable Structure

The probability of retaining $k$ factors was calculated across 1,000 bootstrap iterations. We determined the most stable structure for each configuration by analyzing where the probabilities converged or presented a dominant model.

```{r, fig.height=8, fig.width=20}
retention_p1<-main_results[["SSQ-14 (Original)"]][["retention_analysis"]][["plot"]]
retention_p2<-main_results[["SSQ-10 (Theoretical Depressive Symptoms)"]][["retention_analysis"]][["plot"]]
retention_p3<-main_results[["SSQ-08 (Dixon Chibanda et.al)"]][["retention_analysis"]][["plot"]]

(
  (retention_p1
   )+
  (retention_p2
   )+
  (retention_p3
   )
  ) + 
  patchwork::plot_layout(guides = "collect",
                         axes = "collect_y") +
  patchwork::plot_annotation(title="Factor Retention Probabilities and Counts",
                         tag_level="A"
                  ) &
  
  ggplot2::theme_minimal(base_size = 16,
                         base_family = "Arial") & 
  ggplot2::labs() & 
  ggplot2::theme(
    legend.position = 'bottom',
    plot.caption = ggplot2::element_text(hjust = 0, size = 12, family = "Arial"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    )
```

For the full scale, the Kaiser Criterion strongly favored a 2-factor solution with approximately 89.2% probability. Conversely, Parallel Analysis favored a 3-factor solution with 90.9% probability. These results indicate that the scale is robustly multidimensional. The 2-factor solution represents the most stable base structure, separating somatic from emotional symptoms, while the 3-factor solution represents a stable refinement of the emotional domain.

For the theoretical subset, the Kaiser Criterion overwhelmingly suggested a 1-factor solution in 81.7% of iterations. Parallel Analysis was more variable, favoring three factors in 62.2% of iterations but with significant support for two factors in 35.1% of cases. Given the known sensitivity of PA to minor factors and the strong consensus for one factor under the Kaiser Criterion, the 1-factor structure appears to be the most practical and stable configuration for this subset. The additional factors suggested by PA likely reflect minor correlations rather than substantive clinical dimensions.

For the SSQ-08 subset, the Kaiser Criterion universally suggested a 1-factor solution (100% probability). Parallel Analysis was ambiguous, splitting roughly evenly between 2-factor (37.9%) and 3-factor (37.2%) solutions, with a notable minority (24.4%) supporting a 1-factor solution. Based on these data, this subset is determined to be robustly uni-dimensional. The PA results likely reflect item doublets, or pairs of highly correlated items, rather than distinct latent constructs.

### Structural Stability and Item Co-occurrence

To validate these determinations, we examined the stability of item memberships within the winning models for each method.

#### SSQ-14 (Origina)

##### Kaiser Criterion (2-Factor Solution)

Under the Kaiser criterion, the scale splits cleanly into two distinct clusters. The high probabilities in the stability plot indicate very little cross-loading across iterations. The first factor, labeled General/Somatic, comprises items SSQ09 through SSQ14, which include hallucinations, suicide, and somatic complaints. The second factor, labeled Depressive/Anxious, includes items SSQ01 through SSQ08, encompassing symptoms such as thinking too much, sleep difficulties, and tearfulness. This confirms the scale measures two distinct but related constructs: general distress or somatic symptoms and core depression or anxiety.

```{r, fig.height=10, fig.width=16}
cooccurrence_p<-main_results[["SSQ-14 (Original)"]][["cooccurrence"]][["Kaiser"]][["plot_heatmap"]]
stability_p<-main_results[["SSQ-14 (Original)"]][["cooccurrence"]][["Kaiser"]][["plot_stability"]]

((cooccurrence_p)+(stability_p)) + 
  patchwork::plot_layout() +
  patchwork::plot_annotation(title="Item co-occurrence for the winning factor count and item stability sorted by Winning Model",
                         tag_level="A"
                  ) &
  ggplot2::theme_minimal(base_size = 16) & 
  ggplot2::theme(legend.position = 'bottom',
                 axis.text.x  = ggplot2::element_text(angle = 45,vjust=.85,hjust=1),
                 )
```

```{r}
loadings_kaiser <- extract_loadings_long(main_results, "SSQ-14 (Original)", method = "Kaiser",factor_filter=2)
eigen_scree     <- extract_eigen_scree(main_results, "SSQ-14 (Original)")
```

##### Parallel Analysis (3-Factor Solution)

Parallel Analysis consistently splits the Depressive factor further. The first factor remains the Somatic/Severe dimension, comprising items SSQ07 and SSQ09 through SSQ14. A second factor, labeled Core Depression, includes items SSQ01 through SSQ03. A third factor, labeled Cognitive/Anxiety, includes items SSQ04 through SSQ06 and SSQ08. While stable, this granularity may offer less clinical utility than the 2-factor solution depending on the intended use.

```{r, fig.height=10, fig.width=16}
cooccurrence_p<-main_results[["SSQ-14 (Original)"]][["cooccurrence"]][["Parallel"]][["plot_heatmap"]]
stability_p<-main_results[["SSQ-14 (Original)"]][["cooccurrence"]][["Parallel"]][["plot_stability"]]

((cooccurrence_p)+(stability_p)) + 
  patchwork::plot_layout() +
  patchwork::plot_annotation(title="Item co-occurrence for the winning factor count and item stability sorted by Winning Model",
                         tag_level="A"
                  ) &
  ggplot2::theme_minimal(base_size = 16) & 
  ggplot2::theme(legend.position = 'bottom',
                 axis.text.x  = ggplot2::element_text(angle = 45,vjust=.85,hjust=1),
                 )
```

#### SSQ-10 (Theoretical Subset)

This subset excludes items SSQ04 through SSQ07 to focus on depressive symptoms. Although the Kaiser Criterion supports a 1-factor model, examining the Parallel Analysis 3-factor solution reveals why a 1-factor model might be superior for this subset.

##### Parallel Analysis (3-Factor Solution)

The structure under PA reveals three factors. The first factor includes items SSQ09 through SSQ14, representing severe symptoms. The second factor includes items SSQ01 through SSQ03, representing core emotional symptoms. The third factor consists solely of item SSQ08, representing fatigue or work lag. The emergence of SSQ08 as a singleton or weakly associated factor in the 3-factor solution suggests structural fragmentation. This supports the decision to treat the SSQ-10 as a uni-dimensional scale, as suggested by the Kaiser Criterion, for the purpose of establishing a depression cut-off, as dividing it further creates unstable, item-specific factors.

```{r, fig.height=10, fig.width=16}
cooccurrence_p<-main_results[["SSQ-10 (Theoretical Depressive Symptoms)"]][["cooccurrence"]][["Parallel"]][["plot_heatmap"]]
stability_p<-main_results[["SSQ-10 (Theoretical Depressive Symptoms)"]][["cooccurrence"]][["Parallel"]][["plot_stability"]]

((cooccurrence_p)+(stability_p)) + 
  patchwork::plot_layout() +
  patchwork::plot_annotation(title="Item co-occurrence for the winning factor count and item stability sorted by Winning Model",
                         tag_level="A"
                  ) &
  ggplot2::theme_minimal(base_size = 16) & 
  ggplot2::theme(legend.position = 'bottom',
                 axis.text.x  = ggplot2::element_text(angle = 45,vjust=.85,hjust=1),
                 )
```

#### SSQ-08 (Dixon-Chibanda Subset)

Similarly, while the Kaiser Criterion identifies one factor, the Parallel Analysis 2-factor solution demonstrates instability.

##### Parallel Analysis (2-Factor Solution)

The structure under the PA 2-factor model splits the items into two groups. Factor 1 contains items SSQ11 through SSQ14. Factor 2 contains items SSQ01, SSQ08, SSQ09, and SSQ10. The co-occurrence heatmap shows fuzziness for Factor 2 items, indicating they do not cluster together as reliably as Factor 1 items. This lack of cohesion under the PA 2-factor model reinforces the conclusion that the 1-factor model identified by the Kaiser Criterion is the most accurate representation of the SSQ-08 structure.

```{r, fig.height=10, fig.width=16}
cooccurrence_p<-main_results[["SSQ-08 (Dixon Chibanda et.al)"]][["cooccurrence"]][["Parallel"]][["plot_heatmap"]]
stability_p<-main_results[["SSQ-08 (Dixon Chibanda et.al)"]][["cooccurrence"]][["Parallel"]][["plot_stability"]]

((cooccurrence_p)+(stability_p)) + 
  patchwork::plot_layout() +
  patchwork::plot_annotation(title="Item co-occurrence for the winning factor count and item stability sorted by Winning Model",
                         tag_level="A"
                  ) &
  ggplot2::theme_minimal(base_size = 16) & 
  ggplot2::theme(legend.position = 'bottom',
                 axis.text.x  = ggplot2::element_text(angle = 45,vjust=.85,hjust=1),
                 )
```

#### PHQ-09 (Original)

##### Parallel Analysis (3-Factor Solution)

Although Kaiser strongly supports a 1-factor model, the PA 3-factor solution (Frequency: 209/374 valid runs) reveals a coherent tripartite substructure. Factor 1 (Cognitive/Severe) includes PHQ906–909 (Guilt, Concentration, Motor, Suicide). Factor 2 (Somatic) includes PHQ903–905 (Sleep, Energy, Appetite). Factor 3 (Core Affective) includes PHQ901–902 (Anhedonia, Mood). While this substructure is theoretically interpretable, the low overall stability and high rate of model non-convergence in bootstrapping support using the robust 1-factor structure for the primary clinical objective.

```{r, fig.height=10, fig.width=16}
retention_p<-main_results[["PHQ-09 (Original)"]][["retention_analysis"]][["plot"]]
cooccurrence_p<-main_results[["PHQ-09 (Original)"]][["cooccurrence"]][["Parallel"]][["plot_heatmap"]]
stability_p<-main_results[["PHQ-09 (Original)"]][["cooccurrence"]][["Parallel"]][["plot_stability"]]

((retention_p)+(cooccurrence_p)+(stability_p)) + 
  patchwork::plot_layout() +
  patchwork::plot_annotation(title="Factor Retention and Item co-occurrence for the winning factor count and item stability sorted by Winning Model",
                         tag_level="A"
                  ) &
  ggplot2::theme_minimal(base_size = 16) & 
  ggplot2::theme(legend.position = 'bottom',
                 axis.text.x  = ggplot2::element_text(angle = 45,vjust=.85,hjust=1),
                 )
```


Domain,PHQ-9 Item (Concept),SSQ-14 Item (Concept),Harmonisation Potential
Dysphoria,"Feeling down, depressed, hopeless",Tearfulness; Wishing to be dead,High (Direct conceptual link)
Suicidality,Thoughts of hurting self/better off dead,Wishing to be dead,High (Critical anchor)
Sleep,"Trouble falling/staying asleep, sleeping too much",Fail to sleep,High (Somatic anchor)
Fatigue,Feeling tired or having little energy,Tiredness,High (Somatic anchor)
Cognition,Trouble concentrating,Thinking too much (Kufungisisa),Moderate (Kufungisisa is broader)
Appetite,Poor appetite or overeating,Stomach ache / Loss of appetite,Moderate (SSQ focuses on pain/gut)
Anxiety,None (PHQ is depression only),Heart beating fast; Frightened,Low (Unique to SSQ)
Somatic,General (implicitly covered),Heat in body; Heaviness,Low (Unique to SSQ)
Anhedonia,Little interest or pleasure,No direct equivalent,Low (Information loss in SSQ)

\newpage

## Acknowledgements
[Insert Acknowledgements here]

## Financial Support
[Insert Funding Information here]

## Conflict of Interest
Data Availability
